### 2023-03-16 slides/template.tex
added 1-slide

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 2.0 (March 8, 2022)
%
% This template originates from:
% https://www.LaTeXTemplates.com
%
% Author:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 4.0 (https://creativecommons.org/licenses/by-nc-sa/4.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
	11pt, % Set the default font size, options include: 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt
	%t, % Uncomment to vertically align all slide content to the top of the slide, rather than the default centered
	%aspectratio=169, % Uncomment to set the aspect ratio to a 16:9 ratio which matches the aspect ratio of 1080p and 4K screens and projectors
]{beamer}

\graphicspath{{Images/}{./}} % Specifies where to look for included images (trailing slash required)

\usepackage[english,russian]{babel}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule for better rules in tables

\usepackage{fixltx2e}% for \textsubscript
\usepackage[fleqn]{amsmath}
\usepackage{bm}

%----------------------------------------------------------------------------------------
%	SELECT LAYOUT THEME
%----------------------------------------------------------------------------------------

% Beamer comes with a number of default layout themes which change the colors and layouts of slides. Below is a list of all themes available, uncomment each in turn to see what they look like.

\usetheme{Madrid}

%----------------------------------------------------------------------------------------
%	SELECT COLOR THEME
%----------------------------------------------------------------------------------------

% Beamer comes with a number of color themes that can be applied to any layout theme to change its colors. Uncomment each of these in turn to see how they change the colors of your selected layout theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{monarca}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{spruce}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%----------------------------------------------------------------------------------------
%	SELECT FONT THEME & FONTS
%----------------------------------------------------------------------------------------

% Beamer comes with several font themes to easily change the fonts used in various parts of the presentation. Review the comments beside each one to decide if you would like to use it. Note that additional options can be specified for several of these font themes, consult the beamer documentation for more information.

\usefonttheme{default} % Typeset using the default sans serif font
%\usefonttheme{serif} % Typeset using the default serif font (make sure a sans font isn't being set as the default font if you use this option!)
%\usefonttheme{structurebold} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in bold
%\usefonttheme{structureitalicserif} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in italic serif
%\usefonttheme{structuresmallcapsserif} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in small caps serif

%------------------------------------------------

%\usepackage{mathptmx} % Use the Times font for serif text
\usepackage{palatino} % Use the Palatino font for serif text

%\usepackage{helvet} % Use the Helvetica font for sans serif text
\usepackage[default]{opensans} % Use the Open Sans font for sans serif text
%\usepackage[default]{FiraSans} % Use the Fira Sans font for sans serif text
%\usepackage[default]{lato} % Use the Lato font for sans serif text

%----------------------------------------------------------------------------------------
%	SELECT INNER THEME
%----------------------------------------------------------------------------------------

% Inner themes change the styling of internal slide elements, for example: bullet points, blocks, bibliography entries, title pages, theorems, etc. Uncomment each theme in turn to see what changes it makes to your presentation.

%\useinnertheme{default}
\useinnertheme{circles}
%\useinnertheme{rectangles}
%\useinnertheme{rounded}
%\useinnertheme{inmargin}

%----------------------------------------------------------------------------------------
%	SELECT OUTER THEME
%----------------------------------------------------------------------------------------

% Outer themes change the overall layout of slides, such as: header and footer lines, sidebars and slide titles. Uncomment each theme in turn to see what changes it makes to your presentation.

%\useoutertheme{default}
%\useoutertheme{infolines}
%\useoutertheme{miniframes}
%\useoutertheme{smoothbars}
%\useoutertheme{sidebar}
%\useoutertheme{split}
%\useoutertheme{shadow}
%\useoutertheme{tree}
%\useoutertheme{smoothtree}

%\setbeamertemplate{footline} % Uncomment this line to remove the footer line in all slides
%\setbeamertemplate{footline}[page number] % Uncomment this line to replace the footer line in all slides with a simple slide count

%\setbeamertemplate{navigation symbols}{} % Uncomment this line to remove the navigation symbols from the bottom of all slides

%----------------------------------------------------------------------------------------
%	PRESENTATION INFORMATION
%----------------------------------------------------------------------------------------

\title[]{Анализ смещения распределения в задаче контрастного распределения} % The short title in the optional parameter appears at the bottom of every slide, the full title in the main parameter is only on the title page

\author[Лидия Троешестова \and Роман Исаченко]{Лидия Троешестова \and Роман Исаченко} % Presenter name(s), the optional parameter can contain a shortened version to appear on the bottom of every slide, while the main parameter will appear on the title slide

\date[\today]{ \\ \today} % Presentation date or conference/meeting name, the optional parameter can contain a shortened version to appear on the bottom of every slide, while the required parameter value is output to the title slide

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE SLIDE
%----------------------------------------------------------------------------------------

% \begin{frame}
% 	\titlepage % Output the title slide, automatically created using the text entered in the PRESENTATION INFORMATION block above
% \end{frame}

%------------------------------------------------

\subsection{problem}
\setbeamerfont{frametitle}{size=\small}

\begin{frame}
    \frametitle{Анализ смещения распределения в задаче контрастного распределения}
\scriptsize 
$p^+_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как позитивный объект для $\mathbf{x}$.

% = h(\mathbf{x})) p(\mathbf{x'}|h(\mathbf{x'}) = p(\mathbf{x'}|h(\mathbf{x'}) \neq h(\mathbf{x}))

$p^-_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как негативный объект для $\mathbf{x}$.

$\tau^+$ — вероятность 1 класса; $\tau^- = 1 - \tau^+$ — вероятность любого другого класса $\implies p(\mathbf{x'}) = \tau^+ p_x^+(\mathbf{x'}) + \tau^-p_x^-(\mathbf{x'})$

\textbf{Цель}: исследовать методы устранения смещения распределений $p_x^-$ и $p_x^+$ без использования меток соответствующих классов.

 \textbf{Идея}: в новой функции потерь использовать оценку распределения $p_x^+$ в предопложении верности негативных объектов.

\begin{figure}
\centering
\begin{subfigure}
  \centering
  \includegraphics[width=.4\linewidth]{Images/triplet.jpeg}
  % \caption{Triplet loss visualization}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}
  \centering
  \includegraphics[width=.5\linewidth]{Images/t-SNE.jpeg}
  % \caption{t-SNE visualization of learned representations on CIFAR10}
  \label{fig:sub2}
\end{subfigure}
\label{fig:test}
\end{figure}

\begin{equation*} \small
L_{\text{Unbiased}}^N(f) = \mathbb{E}_{\substack{\mathbf{x} \sim p, \mathbf{x^+} \sim p^+_x,\\ \mathbf{x^-} \sim p_x^-}} \bigg[-\log \frac{\exp (\text{sim}_f(\mathbf{x}, \mathbf{x^+}))}{\exp (\text{sim}_f(\mathbf{x}, \mathbf{x^+}) + \sum _{i=1}^N \exp (\text{sim}_f(\mathbf{x}, \mathbf{x^-})}\bigg]

\end{equation*}
\end{frame}

%------------------------------------------------

\end{document}

### 2023-03-27 code/main.py
fixed dropFN plot

@@ -291,6 +291,7 @@ if __name__ == '__main__':
         top_k=args.top_k,
         epochs=args.epochs,
         num_pos=args.num_pos,
+        drop_fn=args.drop_fn,
         noise_frac=args.noise_frac,
         m_agg_mode=args.m_agg_mode,
         lr=args.lr,

### 2023-04-06 code/loss.py
theory,+ full document

@@ -89,9 +89,9 @@ class DebiasedNegLoss(ContrastiveLossBase):
 
         pos_m = [pos]
         for vec in out_m:
-            pos_1 = torch.exp(torch.sum(out_1 * vec, dim=-1) / self.temperature)
-            pos_2 = torch.exp(torch.sum(out_2 * vec, dim=-1) / self.temperature)
-            pos_new = torch.cat([pos_1, pos_2], dim=0)
+            pos_1 = torch.exp(torch.sum(out_1 * vec, dim=-1) / self.temperature)  # shape (bs)
+            pos_2 = torch.exp(torch.sum(out_2 * vec, dim=-1) / self.temperature) # shape (bs)
+            pos_new = torch.cat([pos_1, pos_2], dim=0)  # shape (2 * bs)
             pos_m.append(pos_new)
         pos_m = torch.stack(pos_m, dim=0).mean(dim=0)
 
@@ -178,6 +178,53 @@ class DebiasedPosLossV2(nn.Module):
         return "Temperature: {}\nCuda: {}\nDrop FN{}\nTau plus: {}".format(
             self.temperature, self.cuda, self.drop_fn, self.tau_plus)
 
+class DebiasedPosLossV3(nn.Module):
+    def __init__(self, temperature, cuda, drop_fn, tau_plus):
+        super().__init__()
+        self.temperature: float = temperature
+        self.cuda: bool = cuda
+        self.drop_fn: bool = drop_fn
+        self.tau_plus: float = tau_plus
+
+    def forward(self, out_1, out_2, out_m, target):
+        batch_size = out_1.shape[0]
+
+        # estimator g()
+        N = batch_size * 2 - 2
+        # neg score
+        out = torch.cat([out_1, out_2], dim=0)  # shape (2 * bs, fdim)
+        # скалярное произведение всех пар
+        neg = torch.exp(torch.mm(out, out.t().contiguous()) / self.temperature)  # shape (2 * bs, 2 * bs)
+        p_estimate = neg.sum(dim=-1) / (N + 2) # shape (2 * bs)
+        # Drop false-negaive pairs
+        if self.drop_fn:
+            mask = get_target_mask(target)
+            if self.cuda:
+                mask = mask.cuda()
+            neg = neg.masked_fill(mask, 0.0)
+
+        # 1ая часть (2 * bs, bs) соответствует одной картинке, 2ая часть (2 * bs, bs+1 - 2 * bs) - другой
[…]

### 2023-04-06 slides/template.tex
middle talk

@@
 
-\usepackage[english,russian]{babel}
 \usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule for better rules in tables
-
-\usepackage{fixltx2e}% for \textsubscript
-\usepackage[fleqn]{amsmath}
-\usepackage{bm}
+\usepackage[russian]{babel}
+% \usepackage[english,russian]{babel}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{xcolor}
 
@@
 
+%\usetheme{default}
+%\usetheme{AnnArbor}
+%\usetheme{Antibes}
+%\usetheme{Bergen}
+%\usetheme{Berkeley}
+%\usetheme{Berlin}
+%\usetheme{Boadilla}
+%\usetheme{CambridgeUS}
+%\usetheme{Copenhagen}
+%\usetheme{Darmstadt}
+%\usetheme{Dresden}
+%\usetheme{Frankfurt}
+%\usetheme{Goettingen}
+%\usetheme{Hannover}
+%\usetheme{Ilmenau}
+%\usetheme{JuanLesPins}
+%\usetheme{Luebeck}
 \usetheme{Madrid}
+%\usetheme{Malmoe}
+%\usetheme{Marburg}
+%\usetheme{Montpellier}
+%\usetheme{PaloAlto}
+%\usetheme{Pittsburgh}
+%\usetheme{Rochester}
+%\usetheme{Singapore}
+%\usetheme{Szeged}
+%\usetheme{Warsaw}
 
@@
 %\usepackage[default]{lato} % Use the Lato font for sans serif text
-
+\usepackage{mathrsfs}
+\usepackage{amsmath}
 %----------------------------------------------------------------------------------------
@@
 
-\title[]{Анализ смещения распределения в задаче контрастного распределения} % The short title in the optional parameter appears at the bottom of every slide, the full title in the main parameter is only on the title page
+\title[]{Анализ смещения распределений в контрастивном обучении} % The short title in the optional parameter appears at the bottom of every slide, the full title in the main parameter is only on the title page
 
@@
 
-\date[\today]{ \\ \today} % Presentation date or conference/meeting name, the optional parameter can contain a shortened version to appear on the bottom of every slide, while the required parameter value is output to the title slide
+\institute[]{МФТИ} % Your institution, the optional parameter can be used for the institution shorthand and will appear on the bottom of every slide after author names, while the required parameter is used on the title slide and can include your email address or additional information on separate lines
+
+\date[\today]{\today} % Presentation date or conference/meeting name, the optional parameter can contain a shortened version to appear on the bottom of every slide, while the required parameter value is output to the title slide
 
@@
 
-% \begin{frame}
-% 	\titlepage % Output the title slide, automatically created using the text entered in the PRESENTATION INFORMATION block above
-% \end{frame}
+\begin{frame}
+	\titlepage % Output the title slide, automatically created using the text entered in the PRESENTATION INFORMATION block above
+\end{frame}
 
+%----------------------------------------------------------------------------------------
+%	PRESENTATION BODY SLIDES
+%----------------------------------------------------------------------------------------
+
+\setbeamerfont{frametitle}{size=\normalsize}
 %------------------------------------------------
 
-\subsection{problem}
-\setbeamerfont{frametitle}{size=\small}
+\begin{frame}
+	\frametitle{Цель исследования}
+
+\begin{block}{Цель}
+Исследовать методы устранения смещения распределений $p_x^-$ и $p_x^+$ без использования меток соответствующих классов.
+\end{block}
+
+\begin{block}{Идея}
+В новой функции потерь использовать оценку распределения $p_x^+$ в предопложении верности негативных объектов.
+\end{block}
+
+\begin{equation*} \small
+L_{\text{Unbiased}}^N(f) = \mathbb{E}_{\substack{\mathbf{x} \sim p, \mathbf{x^+} \sim p^+_x,\\ \mathbf{x^-} \sim p_x^-}} \bigg[-\log \frac{\exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^+}))}{\exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^+}) + \sum _{i=1}^N \exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^-})}\bigg]
+\end{equation*}
+
+\end{frame}
+%------------------------------------------------
 
 \begin{frame}
-    \frametitle{Анализ смещения распределения в задаче контрастного распределения}
-\scriptsize 
+    \frametitle{Постановка задачи}
+\small
 $p^+_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как позитивный объект для $\mathbf{x}$.
-
-% = h(\mathbf{x})) p(\mathbf{x'}|h(\mathbf{x'}) = p(\mathbf{x'}|h(\mathbf{x'}) \neq h(\mathbf{x}))
 
@@
 $\tau^+$ — вероятность 1 класса; $\tau^- = 1 - \tau^+$ — вероятность любого другого класса $\implies p(\mathbf{x'}) = \tau^+ p_x^+(\mathbf{x'}) + \tau^-p_x^-(\mathbf{x'})$
-
-\textbf{Цель}: исследовать методы устранения смещения распределений $p_x^-$ и $p_x^+$ без использования меток соответствующих классов.
-
- \textbf{Идея}: в новой функции потерь использовать оценку распределения $p_x^+$ в предопложении верности негативных объектов.
 
@@
 
+Найти $L_{\text{DebiasedPos}}^N$, минимизирующее $\lim \limits_{N \to \infty} \big|L_{\text{DebiasedPos}}^N (f)  - L_{\text{Unbiased}}^N(f)\big|$.
+
 \begin{equation*} \small
-L_{\text{Unbiased}}^N(f) = \mathbb{E}_{\substack{\mathbf{x} \sim p, \mathbf{x^+} \sim p^+_x,\\ \mathbf{x^-} \sim p_x^-}} \bigg[-\log \frac{\exp (\text{sim}_f(\mathbf{x}, \mathbf{x^+}))}{\exp (\text{sim}_f(\mathbf{x}, \mathbf{x^+}) + \sum _{i=1}^N \exp (\text{sim}_f(\mathbf{x}, \mathbf{x^-})}\bigg]
+L_{\text{Unbiased}}^N(f) = \mathbb{E}_{\substack{\mathbf{x} \sim p, \mathbf{x^+} \sim p^+_x,\\ \mathbf{x^-} \sim p_x^-}} \bigg[-\log \frac{\exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^+}))}{\exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^+}) + \sum _{i=1}^N \exp (\scriptsize\text{sim}_f(\mathbf{x}, \mathbf{x^-})}\bigg]
+\end{equation*}
 
+\end{frame}
+%------------------------------------------------
+
+\begin{frame} 
+	\frametitle{Список литературы}
+
+	\begin{thebibliography}{99} 
+		\footnotesize
+
+		\bibitem[Chuang, 2020]{p1}
+			Ching-Yao Chuang, Joshua Robinson, Lin Yen-Chen, Antonio Torralba, Stefanie Jegelkah (2020)
+			\newblock Debiased Contrastive Learning
+
+            \bibitem[khosla2021supervised, 2020]{p2}
+			Prannay Khosla, Piotr Teterwak, Chen Wang, Aaron Sarna, Yonglong Tian, Phillip Isola, Aaron Maschinot, Ce Liu, Dilip Krishnan (2021)
+			\newblock Supervised Contrastive Learning
+
+           \bibitem[Chen2020SimCLR, 2020]{p3}
+			CTing Chen, Simon Kornblith, Mohammad Norouzi, Geoffrey Hinton (2020)
+			\newblock A Simple Framework for Contrastive Learning of Visual Representations
+
+   		\bibitem[Sohn, 2016]{p4}
+			Sohn, Kihyuk (2016)
+			\newblock Improved Deep Metric Learning with Multi-class N-pair Loss Objective
+
+   		\bibitem[schroff2015facenet, 2015]{p5}
+			Florian Schroff, Dmitry Kalenichenko, James Philbin (2015)
+			\newblock FaceNet: A Unified Embedding for Face Recognition and Clustering
+
+	\end{thebibliography}
+\end{frame}
+%------------------------------------------------
+
+\begin{frame}
+    \frametitle{Решение}
+\scriptsize
+Обозначим $h(\textbf{x}, \widetilde{\textbf{x}}) := e^{f(\textbf{x})^T f(\widetilde{\textbf{x}})}$.
+
+\begin{lemma}
+При $N \to \infty$:
+\begin{equation*}
+L_{\text{Unbiased}}^N(f) 
+\longrightarrow
+\mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^- \sim p_x^-}} \bigg[ - \log \frac{R}{R + N \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}^-)}\bigg],
 \end{equation*}
+где
+\begin{equation*}
+R = \frac{1}{\tau^+} \big(\color{violet} \mathbb{E}_{\textbf{x}' \sim p} h(\textbf{x}, \textbf{x}') \color{black}  - \tau^- \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}^-)\big).
+\end{equation*}
+\end{lemma}
+
+\begin{equation*}
+\tilde{L}_{\text{DebiasedPos}}^N (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^- \sim p_x^-}} \bigg[ - \log \frac{\color{violet} \mathbb{E}_{\textbf{x}' \sim p} h(\textbf{x}, \textbf{x}') \color{black} - \tau^- \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}_i^-)}{\color{violet} \mathbb{E}_{\textbf{x}' \sim p} h(\textbf{x}, \textbf{x}') \color{black} + \big(N \tau^+ - \tau^-\big) \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}_i^-)}\bigg]
+\end{equation*}
+
+Оценим неизвестные матожидания эмпирически:
+\begin{equation*}
+\color{violet}P_{\text{emp}} (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \textbf{v}) = \frac{1}{N+2} \bigg(\sum \limits_{i=1}^N h(\textbf{x}, \textbf{u}_i) + h(\textbf{x}, \textbf{v}) + h(\textbf{x}, \textbf{x})\bigg)\color{black}; \color{blue} P_{\text{emp}}^- (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N) = \frac{1}{N} \sum \limits_{i=1}^N h(\textbf{x}, \textbf{u}_i).
+\end{equation*}
+
+\end{frame}
+%------------------------------------------------
+
+\begin{frame}
+    \frametitle{Решение}
+\scriptsize
+
+\begin{equation*}
+\tilde{L}_{\text{DebiasedPos}}^N (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^- \sim p_x^-}} \bigg[ - \log \frac{\color{violet} \mathbb{E}_{\textbf{x}' \sim p} h(\textbf{x}, \textbf{x}') \color{black} - \tau^- \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}_i^-)}{\color{violet} \mathbb{E}_{\textbf{x}' \sim p} h(\textbf{x}, \textbf{x}') \color{black} + \big(N \tau^+ - \tau^-\big) \color{blue} \mathbb{E}_{\textbf{x}^- \sim p_x^-} h(\textbf{x}, \textbf{x}_i^-)}\bigg]
+\end{equation*}
+
+Финальная оценка:
+\begin{equation*}
+L_{\text{DebiasedPos}}^N (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \{\textbf{u}_i\}_{i=1}^N \sim {p_x^-}^N \\ \textbf{v} \sim p_x^+}} \bigg[-\log \frac{\color{violet}P_{\text{emp}} \color{black} - \tau^- \color{blue} P_{\text{emp}}^-} {\color{violet}P_{\text{emp}} \color{black} + \big(N \tau^+ - \tau^-\big) \color{blue} P_{\text{emp}}^- }\bigg].
+\end{equation*}
+
+\begin{theorem}
+Для произвольного эмбеддинга f и произвольного $\delta > 0$ существует достаточно большое $N$, что
+\begin{equation*}
+\big|\tilde{L}_{\text{DebiasedPos}}^N (f) - L_{\text{DebiasedPos}}^N (f)\big| \leq \bigg[\bigg(1 + \frac{\tau^-}{\tau^+} + \delta\bigg) \sqrt{\frac{\pi}{2N}} + \bigg(1 + \frac{1}{\tau^+}\bigg) \sqrt{\frac{\pi}{2N + 2}}\bigg] e^{3/2}
+\end{equation*}
+\end{theorem}
+
+\end{frame}
+%------------------------------------------------
+
+% \begin{frame}
+% 	\frametitle{Debiased Contrastive Learning}
+%  \framesubtitle{Ching-Yao Chuang et al., NeurIPS 2020} % Optional subtitle
+
+% \scriptsize Минимизация $L_{\text{Debiased}}$ соответствует минимизации верхней границы $L_{\text{Unbiased}}$ в задаче обучения с учителем.
+
+% \begin{equation*} \tiny
+% L_{\text{Debiased}}^{N, M} (f) = \mathbb{E}_{\substack{\mathbf{x} \sim p; \mathbf{x}_+ \sim p_x^+,\\ \{\mathbf{u_i}\}_{i=1}^N \sim p^N \\ \{\mathbf{v_i}\}_{i=1}^M \sim p_x^+^M}}  \bigg[ -\log \frac{e^{f(\mathbf{x})^T f(\mathbf{x}^+)}}{e^{f(\mathbf{x})^T f(\mathbf{x}^+)} + N g\big(\mathbf{x}, \{\mathbf{u_i}\}_{i=1}^N, \{\mathbf{v_i}\}_{i=1}^M\big)} \bigg],
+% \end{equation*}
+
+% \begin{equation*} \tiny 
+% g\bigg(\mathbf{x}, \{\mathbf{u_i}\}_{i=1}^N, \{\mathbf{v_i}\}_{i=1}^M\bigg) = \max \bigg\{ \frac{1}{\tau^-}\bigg(\frac{1}{N} \sum \limits_{i=1}^N e^{f(\mathbf{x})^T f(\mathbf{u_i})} - \tau^+ \frac{1}{M} \sum \limits_{i=1}^M e^{f(\mathbf{x})^T f(\mathbf{v_i})}\bigg), e^{-1/t}\bigg\}
+% \end{equation*}
+
+% \begin{figure}
+%     \includegraphics[width=0.7\linewidth]{Images/tsne.png}
+% \end{figure}
+
+% \end{frame}
+
+%------------------------------------------------
+
+\begin{frame}
+	\frametitle{Использование SimCLR}
+	\framesubtitle{CTing Chenet al., 2020} % Optional subtitle
+
+\begin{figure}
+    \includegraphics[width=0.5\linewidth]{Images/simclr.png}
+\end{figure}
+
+\scriptsize
+\begin{itemize}
+    \item $\mathcal{T} -$ семейство аугментаций (color distortion, gaussian blur) 
+    \item Семплируются 2 аугментации $t, t' \sim \mathcal{T}$, применяются к каждому объекту.
+    \item Обучаем сеть-энкодер $f(\cdot)$ и MLP сеть-проекцию $g(\cdot)$, максимизируя соответсвтие представлений.
+\end{itemize}
 \end{frame}
@@
 
+\begin{frame}
+	\frametitle{Эксперимент с удалением FN и с добавлением FP}
+\begin{figure}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{Images/base_vs_dropfn.png}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}
+  \centering
+  \includegraphics[width=0.95\linewidth]{Images/base_vs_noise.png}
+  \label{fig:sub2}
+\end{subfigure}
+\label{fig:test}
+\end{figure}
+
+\end{frame}
+%------------------------------------------------
+
+\begin{frame}
+	\frametitle{Эксперимент с увеличением кол-ва положительных семплов}
+\begin{figure}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=1\linewidth]{Images/base_vs_M=2.png}
+  \label{fig:sub1}
+\end{subfigure}
+\label{fig:test}
+\end{figure}
+
+\end{frame}
+%------------------------------------------------
+
+\begin{frame}
+	\frametitle{Заключение}
+
+\scriptisize
+
+\begin{itemize}
+    \item Лосс с устраненным FP-смещением добивается точности, сопоставимой с DebiasedNeg, и имеет превосходство на ранних эпохах.
+    \item В случае удаления ошибок FN функции потерь DebiasedPos и DebiasedNeg имеют сопоставимые точности.
+    \item При сильном зашумлении датасета, когда FP-rate увеличен, DebiasedPos имеет большое преимущество в точности над DebiasedNeg.
+    \item При увеличенном кол-ве положитлеьных семплов точность для всех рассмотренных функций потерь возрастает.
+\end{itemize}
+
+\end{frame}
+
+%------------------------------------------------------------------------------
+%	CLOSING SLIDE
+%------------------------------------------------------------------------------
+
+% \begin{frame}[plain] % The optional argument 'plain' hides the headline and footline
+% 	\begin{center}	
+% 		\bigskip\bigskip % Vertical whitespace
+% 		{\LARGE Спасибо за внимание!}
+% 	\end{center}
+% \end{frame}
+
+%----------------------------------------------------------------------------------------
+
 \end{document}

### 2023-04-12 code/loss.py
code changes

@@ -28,6 +28,19 @@ def get_target_mask(target: torch.Tensor) -> torch.Tensor:
     return mask
 
 
+def replace_pos_diagonals(full_matrix, pos_m):
+    """
+    Replace positive diagonals in full_matrix shaped (2bs, 2bs) by pos_m vector
+    """
+    batch_size = full_matrix.shape[0] // 2
+
+    new_matrix = full_matrix.clone()
+
+    new_matrix[range(batch_size), range(batch_size, 2*batch_size)] = pos_m[:batch_size]
+    new_matrix[range(batch_size, 2*batch_size), range(batch_size)] = pos_m[batch_size:]
+    return new_matrix
+
+
 class ContrastiveLossBase(nn.Module):
     def __init__(self, temperature, cuda, drop_fn):
         super().__init__()
@@ -178,6 +191,7 @@ class DebiasedPosLossV2(nn.Module):
         return "Temperature: {}\nCuda: {}\nDrop FN{}\nTau plus: {}".format(
             self.temperature, self.cuda, self.drop_fn, self.tau_plus)
 
+
 class DebiasedPosLossV3(nn.Module):
     def __init__(self, temperature, cuda, drop_fn, tau_plus):
         super().__init__()
@@ -187,36 +201,44 @@ class DebiasedPosLossV3(nn.Module):
         self.tau_plus: float = tau_plus
 
     def forward(self, out_1, out_2, out_m, target):
+        ######################## from super
         batch_size = out_1.shape[0]
 
-        # estimator g()
-        N = batch_size * 2 - 2
-        # neg score
-        out = torch.cat([out_1, out_2], dim=0)  # shape (2 * bs, fdim)
-        # скалярное произведение всех пар
-        neg = torch.exp(torch.mm(out, out.t().contiguous()) / self.temperature)  # shape (2 * bs, 2 * bs)
-        p_estimate = neg.sum(dim=-1) / (N + 2) # shape (2 * bs)
-        # Drop false-negaive pairs
+        out = torch.cat([out_1, out_2], dim=0)  # (2 * bs, fdim)
+        full = torch.exp(torch.mm(out, out.t().contiguous()) / self.temperature)  # (2 * bs, 2 * bs)
         if self.drop_fn:
             mask = get_target_mask(target)
             if self.cuda:
                 mask = mask.cuda()
-            neg = neg.masked_fill(mask, 0.0)
[…]

### 2023-04-12 code/main.py

@@ -175,8 +175,8 @@ def main(dataset: str, loss: str, root: str, batch_size: int, model_arch, *, cud
     wandb.config.update(config)
     logger.info("\nStart experiment with config: %s\n", config)
     m_agg_mode = utils.MAggMode[m_agg_mode]
-    if m_agg_mode == utils.MAggMode.pos_grouping and loss != "DebiasedNeg" and num_pos > 1:
-        raise Exception("Mean aggregation only available in DebiasedNeg loss")
+    # if m_agg_mode == utils.MAggMode.pos_grouping and loss != "DebiasedNeg" and num_pos > 1:
+    #     raise Exception("Mean aggregation only available in DebiasedNeg loss")
 
     train_loader = DataLoader(
         get_dataset(dataset, root=root, split="train+unlabeled", num_pos=num_pos,

### 2023-04-12 code/loss.py
beautified code

@@ -7,8 +7,9 @@ import torch.nn as nn
 
 def get_negative_mask(batch_size):
     """
-    Дла каждого изображения отмечается исходная картинка и ее аугментация
-    Получается симметричная матрица с нулевой главной диагональю и 2мя параллельными
+    For both images, takes the positions corresponding to itself and
+    to its augmentation. Resulting mask is a matrix (2bs, 2bs) with 
+    the main diagonal and 2 positive parallel sub-diagonals zeroed-out
     """
     negative_mask = torch.ones((batch_size, 2 * batch_size), dtype=bool)
     for i in range(batch_size):
@@ -23,14 +24,14 @@ def get_target_mask(target: torch.Tensor) -> torch.Tensor:
     """
     Generate bool matrix for equal targets
     """
-    target_ = torch.cat([target, target], dim=0).view(-1, 1)  # shape (2 * bs, 1)
-    mask = (target_ == target_.t().contiguous()) & (target_ != -1)  # shape (2 * bs, 2 * bs)
+    target_ = torch.cat([target, target], dim=0).view(-1, 1)  # (2bs,)
+    mask = (target_ == target_.t().contiguous()) & (target_ != -1)  # (2bs, 2bs)
     return mask
 
 
 def replace_pos_diagonals(full_matrix, pos_m):
     """
-    Replace positive diagonals in full_matrix shaped (2bs, 2bs) by pos_m vector
+    Replace positive sub-diagonals in full_matrix shaped (2bs, 2bs) by pos_m vector
     """
     batch_size = full_matrix.shape[0] // 2
 
@@ -51,29 +52,27 @@ class ContrastiveLossBase(nn.Module):
     def forward(self, out_1, out_2, target):
         batch_size = out_1.shape[0]
 
-        # neg score
-        out = torch.cat([out_1, out_2], dim=0)  # shape (2 * bs, fdim)
-        # скалярное произведение всех пар
-        neg = torch.exp(torch.mm(out, out.t().contiguous()) / self.temperature)  # shape (2 * bs, 2 * bs)
-        # Drop false-negaive pairs
+        ### neg score ###
+        out = torch.cat([out_1, out_2], dim=0)  # (2bs, d)
+        # scalar product of all pairs
+        full = torch.exp(torch.mm(out, out.t().contiguous()) / self.temperature)  # (2bs, 2bs)
[…]

### 2023-04-12 code/main.py

@@ -31,20 +31,20 @@ def train(net, data_loader, loss_criterion, train_optimizer, batch_size, *, cuda
             pos_1, pos_2 = pos_1.cuda(non_blocking=True), pos_2.cuda(non_blocking=True)
             pos_m = [pos.cuda(non_blocking=True) for pos in pos_m]
         t_extract_image = time.time()
-        feature_1, out_1 = net(pos_1)
-        feature_2, out_2 = net(pos_2)
+        _, out_1 = net(pos_1)
+        _, out_2 = net(pos_2)
 
         if m_agg_mode == utils.MAggMode.pos_grouping:
             out_m = []
             for pos_i in pos_m:
-                feature_i, out_i = net(pos_i)
+                _, out_i = net(pos_i)
                 out_m.append(out_i)
             t_forward_pass = time.time()
             loss = loss_criterion(out_1, out_2, out_m, target)
         elif m_agg_mode == utils.MAggMode.loss_combination:
             out_m = [out_1, out_2]
             for pos_i in pos_m:
-                feature_i, out_i = net(pos_i)
+                _, out_i = net(pos_i)
                 out_m.append(out_i)
             t_forward_pass = time.time()
             losses = []
@@ -83,8 +83,6 @@ def train(net, data_loader, loss_criterion, train_optimizer, batch_size, *, cuda
 
         total_num += batch_size
         total_loss += loss.item() * batch_size
-
-        # train_bar.set_description("Train Epoch: [{}/{}] Loss: {:.4f}".format(epoch, epochs, total_loss / total_num))
         t_start = time.time()
     return total_loss / total_num, step
 
@@ -98,7 +96,7 @@ def test(net, memory_data_loader, test_data_loader, *, top_k, class_cnt, cuda=Tr
         for data, _, _, target in memory_data_loader:
             if cuda:
                 data = data.cuda(non_blocking=True)
-            feature, out = net(data)
+            feature, _ = net(data)
             feature_bank.append(feature)
         # [D, N]
         feature_bank = torch.cat(feature_bank, dim=0).t().contiguous()
@@ -112,7 +110,7 @@ def test(net, memory_data_loader, test_data_loader, *, top_k, class_cnt, cuda=Tr
             if cuda:
                 data, target = data.cuda(
                     non_blocking=True), target.cuda(non_blocking=True)
-            feature, out = net(data)
+            feature, _ = net(data)
 
[…]

### 2023-04-12 paper/main.tex
raw final paper

\documentclass{article}
\usepackage{arxiv}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{enumitem,kantlipsum}
\usepackage{mathtools}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage{hyperref}
\usepackage{cleveref}

\declaretheorem[name=Theorem, numberwithin=section]{thm}

\title{Mitigating distributional biases in contrastive learning}

\author{ Lidia Troeshestova \\
	Moscow Institute of Physics and Technology\\
	\texttt{troeshestova.ls@phystech.edu} \\
	%% examples of more authors
	\And
	Roman Isachenko\\
	   \\
	\texttt{} \\
	%% \AND
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
}
\date{}

\renewcommand{\shorttitle}{\textit{arXiv} Template}

\hypersetup{
pdftitle={A template for the arxiv style},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={David S.~Hippocampus, Elias D.~Striatum},
pdfkeywords={First keyword, Second keyword, More},
}

\begin{document}
\nocite{*}
\maketitle

\begin{abstract}
	Recently there has been a renewed interest in using contrastive learning for self-supervised representation learning. A popular method for learning representations without labels is to compare similar (positive) and dissimilar (negative) pairs of samples. However, negative samples are often chosen randomly, which means that they could be of the same class. This paper analyzes various ways to eliminate these biases. Based on the fully-supervised case, we develop debiased contrastive models that account for same-label datapoints without requiring knowledge of true labels, and explore their properties. The experiments are performed on CIFAR10 dataset. This will further improve availability of accurate models for classification in tasks where extensive labeling is expensive or inaccessible.
\end{abstract}

\keywords{: Contrastive learning \and Representation learning \and Self-supervised learning}

\section{Introduction}
Representation learning has become increasingly popular in recent years due to its ability to learn meaningful representations from large amounts of data, without the need for manual feature engineering. A widely used solution for this problem is contrastive learning, a technique that uses the principle of contrasting samples against each other to learn attributes that are common between data classes and attributes that set apart a data class from another. It encourages the representations of similar pairs $(\textbf{x}, \textbf{x}^+)$ to be close, and those of dissimilar pairs $(\textbf{x}, \textbf{x}^-)$ to be more orthogonal.

The basic metric for supervised similarity is \textit{triplet loss} \citep{schroff2015facenet}, where \textbf{x} is called \textit{anchor}, \textbf{x}$^+$ is a sample with the same label as anchor, and \textbf{x}$^-$ is a sample with the different label than anchor. An improvement of triplet loss is \textit{Multi-class N-pair Loss} invented by \citep{NIPS2016_6b180037}:

\begin{equation} \label{eq:1}
\mathcal{L}(\textbf{x}, \textbf{x}^+, \{\textbf{x}_i^-\}_{i=1}^{N-1}; f) = - \log \frac{\exp(f(\textbf{x})^T f(\textbf{x}_i^+))}{\exp(f(\textbf{x})^T f(\textbf{x}_i^+)) + \sum _{i=1}^{N-1} \exp(f(\textbf{x})^Tf(\textbf{x}_i^-))}
\end{equation}

Contrastive learning has proven to be effective for eliciting visual representations: SimCLR  \citep{Chen2020SimCLR} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods. Its authors developed NT-Xent loss (normalized temperature-scaled cross entropy loss), that uses cosine similarity instead of scalar products.

Due to the unavailability of actual labels during training, negative instances $\textbf{x}^-$ are often randomly selected from the training data. However, this approach results in sampling bias, where $\textbf{x}^-$ may actually be similar to $\textbf{x}$. This bias causes a significant drop in performance.

To address this issue, \citep{khosla2021supervised} extend the self-supervised batch contrastive approach to the fully-supervised setting, allowing to effectively leverage label information. According to the experiments, supervised contrastive loss outperforms the base cross entropy, but only by a small amount. It also outperforms the cross entropy on robustness benchmark (ImageNet-C, which applies common naturally occuring perturbations such as noise, blur and contrast changes to the ImageNet dataset) and is less sensitive to hyperparameter changes.

As for the case of self-supervised learning, a debiased contrastive loss was proposed to mitigate the bias of negative sampling by \citep{chuang2021debiased}. Given $N$ negative and $M$ positive samples, $\mathcal{L}^{N, M}_{\text{Debiased}} (f)$ is the loss function that corrects for False Negative samples. To remain unsupervised in practice, samples are acquired from the data distribution and a positive distribution mimicked by data augmentations, which can contatin False Positives.

This work aims to develop a new algorithm for debiased sampling in contrastive learning, which accounts for False Positive as well as False Negative samples.

% In addition, it is suggested to tackle the issue of choosing a wrong anchor in multimodal applications such as attribute detection.

\section{Problem Statement}
\label{sec:headings}

\subsection{Debiased Contrastive Learning}

In \citep{chuang2021debiased} the distribution over classes is denoted by $\rho(c)$, the joint distribution $p_{x,c}(\textbf{x}, c) = p(\textbf{x}|c)\rho(c)$. Let $h : X \rightarrow C$ be the function assigning the latent class labels. Then $p^+_x(\textbf{x}') = p(\textbf{x}'|h(\textbf{x}') = h(\textbf{x}))$ is the probability of observing $\textbf{x}'$ as a positive example for $\textbf{x}$ and $p^-_x(\textbf{x}') = p(\textbf{x}'|h(\textbf{x}') \neq h(\textbf{x}))$ the probability of a negative example. It is assumed that the class probabilities $\rho(\textbf{x}) = \tau^+$ are uniform, and $\tau^- = 1 - \tau^+$ is the probability of observing any different class. Then the  ``ideal``, or unbiased loss to optimize would be:

\begin{equation}  \label{eq:2}
L_{\text{Unbiased}}^N(f) = \mathbb{E}_{\substack{\textbf{x} \sim p, \textbf{x}^+ \sim p^+_x,\\ \textbf{x}_i^- \sim p_x^-}} \bigg[-\log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)}}{e^{f(\textbf{x})^T f(\textbf{x}^+)} + \sum _{i=1}^N e^{f(\textbf{x})^T f(\textbf{x}_i^-)}}\bigg]
\end{equation}

In practice, $p_x^-(\textbf{x}_i^-)$ is unavailable. Thus, the standard solution is to sample negative examples $\textbf{x}_i^-$ from the unlabeled $p(\textbf{x})$ instead. When drawn from $p(\textbf{x})$, the sample $\textbf{x}_i^-$ will be of the same class as $\textbf{x}$ with probability $\tau^+$.

\begin{equation} \label{eq:3}
L_{\text{Biased}}^N(f) = \mathbb{E}_{\substack{\textbf{x} \sim p, \textbf{x}^+ \sim p^+_x,\\ \textbf{x}_i^- \sim p_x}} \bigg[-\log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)}}{e^{f(\textbf{x})^T f(\textbf{x}^+)} + \sum _{i=1}^N e^{f(\textbf{x})^T f(\textbf{x}_i^-)}}\bigg]
\end{equation}

\paragraph{\underline{Problem statement:}} Derive a loss function that approximates the ideal $L_{\text{Unbiased}}^N$, only having access to $M$ positive samples and samples from the distribution $p$.

The data distribution can be decomposed:
\begin{equation} \label{eq:4}
p(\textbf{x}') = \tau^+ p^+_x(\textbf{x}') + \tau^-p_x^-(\textbf{x}')
\end{equation}

And the missing $p^-_x(\textbf{x}')$ is derived as
\begin{equation} \label{eq:5}
p_x^-(\textbf{x}') = \frac{p(\textbf{x}') - \tau^+ p^+_x(\textbf{x}')}{\tau^-}
\end{equation}
However, estimating $p(\textbf{x}')$ and $p^+_x(\textbf{x}')$ empirically is computationally expensive, and the authors propose a cheaper way to estimate $p^-_x(\textbf{x}')$, using $N$ samples $\{\textbf{u}_i\}_{i=1}^N$ from $p$ and $M$ samples $\{\textbf{v}_i\}_{i=1}^M$ from $p_x^+$, with temperature $t$:

\begin{equation} \label{eq:6}
L_{\text{DebiasedNeg}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p; \textbf{x}_+ \sim p_x^+,\\ \{\textbf{u}_i\}_{i=1}^N \sim p^N \\ \{\textbf{v}_i\}_{i=1}^N \sim {p_x^+}^M}}  \bigg[ -\log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)} }{e^{f(\textbf{x})^T f(\textbf{x}^+)} + N g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_i\}_{i=1}^M\big)} \bigg],
\end{equation}

where the empirical estimates of $p$ and $p_x^-$ are defined as
\begin{equation} \label{eq:7}
g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_i\}_{i=1}^M\big) = \max \bigg\{ \frac{1}{\tau^-}\bigg(\frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} - \tau^+ \frac{1}{M} \sum \limits_{i=1}^M e^{f(\textbf{x})^T f(\textbf{v}_i)}\bigg), e^{-1/t}\bigg\}.
\end{equation}

Using this loss on SimCLR framework allowed to achieve higher accuracy compared to the regular biased loss. With a similar theoretical inference, this work is aimed at developing a new algorithm, that would additionally focus on eliminating bias caused by False Positives.

% \paragraph{\underline{Tasks:}} 
% \begin{itemize}
%   \item reproduce Debiased Contrastive Learning results
%   \item develop novel debiased contrastive loss that accounts for False Positives.
%   \item test the algorithm on CIFAR10 and MNIST datasets
% \end{itemize}

\subsection{False Positive Samples} \label{FP}
False Positive samples can be encountered in contrastive learning due to excessive or inappropriate augmentation. Similar to Debiased Contrastive Learning, we propose to estimate the positive sampling distribution in the loss function.

\begin{equation}  \label{eq:8}
p_x^+ (\textbf{x}') = \frac{p(\textbf{x}') - \tau^- p^-_x(\textbf{x}')}{\tau^+}
\end{equation}

Contrary to Debiased Contrastive Learning, drawing samples from $p(\textbf{x}')$ will also include positive labelled datapoints, since some of them could be actually negative.

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\begin{lemma}
% \normalfont 
With $N \to \infty$:
\begin{equation} \label{eq:9}
\mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^+ \sim p_x^+ \\ \{\textbf{x}_i^-\}_{i=1}^N \sim {p_x^-}^N}} \bigg[ - \log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)}}{e^{f(\textbf{x})^T f(\textbf{x}^+)} + \sum_{i=1}^N e^{f(\textbf{x})^T f(\textbf{x}_i^-)}} \bigg] \longrightarrow
\mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^- \sim p_x^-}} \bigg[ - \log \frac{R}{R + N \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}^-)}} \bigg],
\end{equation}

where
\begin{equation}  \label{eq:10}
R = \frac{1}{\tau^+} \big(\mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^T f(\textbf{x}')} - \tau^- \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}^-)}\big).
\end{equation}
\end{lemma}

\renewcommand\qedsymbol{$\blacksquare$}
\begin{proof}
Since the expression inside the expectation is bounded, we can apply Dominated convergence theorem:

\begin{flalign*}
&\lim_{N \to \infty} \mathbb{E} \bigg[ - \log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)}}{e^{f(\textbf{x})^T f(\textbf{x}^+)} + \sum_{i=1}^N e^{f(\textbf{x})^T f(\textbf{x}_i^-)}} \bigg] = \mathbb{E} \bigg[ \lim_{N \to \infty} - \log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)}}{e^{f(\textbf{x})^T f(\textbf{x}^+)} + \sum_{i=1}^N e^{f(\textbf{x})^T f(\textbf{x}_i^-)}} \bigg] = &\\
& = \mathbb{E} \bigg[ - \log \frac{\mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^T f(\textbf{x}^+)}}{\mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^T f(\textbf{x}^+)} + N \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}^-)}}\bigg]
\end{flalign*}

Using \ref{eq:8} and linearity of the expectation we complete the proof:
$$\mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^T f(\textbf{x}^+)} = \frac{1}{\tau^+} \big(\mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^T f(\textbf{x}')} - \tau^- \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}^-)}\big) = R.$$
\end{proof}

We will denote the limit by $\tilde{L}_{\text{DebiasedPos}}^N$ and rewrite it, multiplying the nominator and the denominator of the under-logarithm expression by $\tau^+$:

\begin{equation} \label{eq:11}
\tilde{L}_{\text{DebiasedPos}}^N (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \textbf{x}^- \sim p_x^-}} \bigg[ - \log \frac{\mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^T f(\textbf{x}')} - \tau^- \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}_i^-)}}{\mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^T f(\textbf{x}')} + \big(N \tau^+ - \tau^-\big) \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}_i^-)}}\bigg]
\end{equation}

With finite N:
\begin{equation} \label{eq:12}
L_{\text{DebiasedPos}}^N (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \{\textbf{u}_i\}_{i=1}^N \sim {p_x^-}^N \\ \textbf{v} \sim p_x^+}} \bigg[-\log \frac{P_{\text{emp}} (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \textbf{v}) - \tau^- P_{\text{emp}}^- (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N)} {P_{\text{emp}} (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \textbf{v})+ \big(N \tau^+ - \tau^-\big) P_{\text{emp}}^- (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N) }\bigg],
\end{equation}

where
\begin{equation} \label{eq:13}
P_{\text{emp}} (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \textbf{v}) = \frac{1}{N+2} \bigg(\sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} + e^{f(\textbf{x})^T f(\textbf{v})} + e^{f(\textbf{x})^T f(\textbf{x})}\bigg);
\end{equation}

\begin{equation} \label{eq:14}
P_{\text{emp}}^- (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N) = \frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)}.
\end{equation}

When $N$ is finite and $M=1$, theorem \ref{thm:asympt} bounds the error as decreasing with rate $O(N^{-1/2})$.

\begin{restatable}[]{thm}{asympt}
\label{thm:asympt}
If $M = 1$, for any embedding f and any $\delta > 0$ exists large enough N, that
\begin{equation} \label{eq:15}
\big|\tilde{L}_{\text{DebiasedPos}}^N (f) - L_{\text{DebiasedPos}}^N (f)\big| \leq \bigg[\bigg(1 + \frac{\tau^-}{\tau^+} + \delta\bigg) \sqrt{\frac{\pi}{2N}} + \bigg(1 + \frac{1}{\tau^+}\bigg) \sqrt{\frac{\pi}{2N + 2}}\bigg] e^{3/2}
\end{equation}
\end{restatable}

\subsection{Positive Sample Aggregation} \label{pos_aggr_theory}

When the number of positive samples $M > 1$, similar to \citep{khosla2021supervised} we will aggregate losses using the summation over positive samples located outside of the log. In our case this means taking all possible pairs of same-class (positive) elements, and averaging all corresponding $L_{\text{DebiasedPos}}^N (f)$ losses. We will refer to this method as \textit{pos-grouping}.

% (which yields better accuracy for supervised contrastive losses rather than summation inside of the log)

\begin{equation} \label{eq:16}
L_{\text{DebiasedPos}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \{\textbf{u}_i\}_{i=1}^N \sim {p^-}^N \\ \{\textbf{v}_i\}_{i=1}^M\sim {p^+}^M}} \frac{1}{M} \sum\limits^M \bigg[-\log \frac{P_{\text{emp}} - \tau^- P_{\text{emp}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}\bigg]
\end{equation}

Another way to aggregate positive samples is to do so inside the estimates $P_{\text{emp}}$, following the example of \citep{chuang2021debiased} which we will refer to as \textit{loss-combination}. Emperically, the experiments in Section \ref{pos_aggr} show that while \textit{pos-grouping} is computationally less expensive, \textit{loss-combination} yields better accuracy with larger M.

% \subsection{Wrong Anchor}
% This issue arises in downstream applications of contrastive learning such as image attribute detection, e.g. discrepancy between the image (image representation $f$) and its description (description representation $t$) in online shopping. To solve this problem, when calculating the losses for each picture, we propose to weigh the losses by the cosine similarity of the anchor image vector and the description text vector.

% \begin{equation} \label{eq:15}
% L_{\text{Unbiased}}(x, x^+, \{x_i^-\}_{i=1}^{N-1}; f) = - \text{sim} (f^T, t^T) \log \frac{\exp(f^T f_i^+)}{\exp(f^T f_i^+ + \sum_{i=1}^N \exp(f^T f_i^-)}
% \end{equation}

\section{Experiment}
\subsection{Baseline Reproduction}

The goal of the experiment is to compare Multi-class N-pair Loss \ref{eq:1} with Negative-Debiased Loss \ref{eq:6}. We use CIFAR10 \citep{krizhevsky2009learning} dataset, an image recognition dataset for image classification. It consists of 60000 32x32 color images in 10 classes. The baseline negative-debiased model is SimCLR \citep{Chen2020SimCLR} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{Diederik2014Adam}, learning rate 0.001, and batch size 512. All models are trained for 50 epochs and evaluated by training a linear classifier after fixing the learned embedding.

 The primary results are shown in Figure \ref{fig:fig2}. While the difference between Top-5 accuracy for the debiased model and the biased model is insignificant, Top-1 accuracy of the debiased model is higher than that of the biased model.

% \begin{figure}[!htb]
% \minipage{0.49\textwidth}
% \includegraphics[width=\linewidth]{figures/baseline_acc1.png}
% \endminipage\hfill
% \minipage{0.49\textwidth}
% \includegraphics[width=\linewidth]{figures/baseline_acc5.png}
% \endminipage
% \caption{Classification accuracy on CIFAR10. Negative-Debiased loss is denoted as debiasedNeg.}
% \label{fig:fig1}
% \end{figure}

\subsection{Testing New Loss}
We run the same experiment for Positive-Debiased loss [\ref{eq:14}]. The code is available at our \href{https://github.com/intsystems/2023-Project-123}{GitHub}. Figure \ref{fig:fig2} shows a tangible difference between the performances of the novel loss and the previous losses. The accuracy advantage is prevalent from the very first steps of training both for Top-1 and Top-5 accuracy, meaning Positive-Debiased loss is more efficient in the early stages of training.

\begin{figure}[!htb]
\minipage{0.49\textwidth}
\includegraphics[width=\linewidth]{figures/3_losses.png}
\endminipage\hfill
\minipage{0.49\textwidth}
\includegraphics[width=\linewidth]{figures/3_losses_5.png}
\endminipage
\caption{Classification accuracy on CIFAR10, Negative- and Positive-Debiased losses are denoted as debiasedNeg and debiasedPos.}
\label{fig:fig2}
\end{figure}

To test the performance of the loss under the assumptions made in \ref{FP}, we artificially exclude False Negative errors via accessing ground truth labels during training, and discover that Positive-Debiased loss still outperforms Negative-Debiased loss [Figure \ref{fig:fig3}].

\begin{figure}
\includegraphics[width=1\textwidth]{figures/base_vs_dropfn.png}
\caption{Classification accuracy on CIFAR10 with dropping False Negatives}
\label{fig:fig3}
\end{figure}

To understand if Positive-Debiased loss is robust to noise, we apply Gaussian blur on train images with probability 0.3, and thus increase the False Positive rate. Figure \ref{fig:fig4} demonstrates that Positive-Debiased loss is able to mitigate False Positive errors significantly better than Negative-Debiased and Contrastive losses.

\begin{figure}
\includegraphics[width=1\textwidth]{figures/base_vs_noise.png}
\caption{Classification accuracy on CIFAR10 with an increased False Positive rate}
\label{fig:fig4}
\end{figure}

\begin{figure}
\includegraphics[width=1\textwidth]{figures/base_vs_M=2.png}
\caption{Classification accuracy on CIFAR10 with an increased positive sample size M}
\label{fig:fig5}
\end{figure}

Additionally we test all three losses with increased positive sample size M [Figure \ref{fig:fig5}]. Larger M leads to better estimate of the loss both for Positive- and Negative-Debiased losses, yielding better accuracy as a result. It is worth noting that in all of the experiments Positive-Debiased loss achieves better accuracy in the earlier stages of training both for Top-1 and Top-5 scores.

\begin{figure}
\includegraphics[width=1\textwidth]{figures/pg_vs_lc.png}
\caption{Classification accuracy on CIFAR10 with various positive sample aggregations}
\label{fig:fig6}
\end{figure}

\subsection{Positive Sample Aggregation} \label{pos_aggr}
In this experiment we compare methods of aggregation over positive samples introduced in Section \ref{pos_aggr_theory}. We train SimCLR with the same parameters, except for setting batch size to 64 when testing $M=4$ for faster training. As expected, there is a trade-off between accuracy and speed. The results in Table \ref{tab:table} show that \textit{pos-grouping} improves over \textit{loss-combination} in accuracy, but is computationally slower. Figure \ref{fig:fig6} shows that with bigger $M$ the advantage of \textit{loss-combination} over \textit{pos-grouping} gets more significant.

\begin{table}
    \caption{Aggregation over M positive samples for DebiasedPos}
    \centering
    \begin{tabular}{llllll}
        \toprule
        Method  & M & bs    & Training time (min)   & acc@1 & acc@5\\
        \midrule
        -   & 1 & 512   & 122.85    & 76.76 & 98.61\\
        \midrule
        pos-grouping    & 2 & 512   & \textbf{181.87}   & 80.1  & 98.86\\
        loss-combination    & 2 & 512   & 192.82    & \textbf{80.5} & \textbf{98.93}\\
        \midrule
        pos-grouping    & 4 & 64    & \textbf{278.88}   & 81.75 & 99.01\\
        loss-combination    & 4 & 64    & 308.25    & \textbf{82.44} & \textbf{99.25}\\
        \bottomrule
    \end{tabular}
    \label{tab:table}
\end{table}

\subsection{Discussion}
\begin{enumerate}
    \item Positive-Debiased loss is based on the assumption of true negative samples, i.e. with no False Negative errors, however, it still improves accuracy of classification when this assumption does not hold.
    \item Another implicit assumption we make is that the classes have uniform distribution, which is not true for many datasets, and it is an interesting direction of further reseach to additionally account for the biases in the class distribution.
\end{enumerate}

\section{Conclusion}
In this paper, we propose a new method for debiasing contrastive loss based on mitigating biases caused by False Positive errors. Theoretical analysis shows that computationally Positive-Debiased and baseline losses are equivalent, however, in a scenario with an increased False Positive rate our loss yields better accuracy, which shows that it is more robust. Additionally, Positive-Debiased loss is more dominant in the early epochs. We further explore 2 mehtods of aggregating loss over positive samples, and infer that \textit{pos-grouping} trains faster, while \textit{loss-combination} leads to better performance.

% Interesting directions of future work include (1) trying the debiased objective in semi-supervised learning or few shot learning, and (2) studying the effect of how positive (similar) examples are drawn, e.g., analyzing different data augmentation techniques.

% \noindent\rule{16cm}{0.4pt}

\clearpage
\bibliographystyle{unsrtnat}
\bibliography{references}

\appendix
\newpage
\section{Proof of Theorem 2}
We recall \cref{thm:asympt}:

\asympt*

\begin{proof}
Let us denote:
\begin{flalign*}
    &P_{\text{est}} := \mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^Tf(\textbf{x}')}&\\
    &P_{\text{est}}^- := \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^Tf(\textbf{x}^-)}&\\
    &A_1 := -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-}&\\
    &A_2 := -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}&\\
    &A_3 := -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}&\\
    &A_4 := -\log \frac{P_{\text{emp}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}&\\
    &A_5 := -\log \frac{P_{\text{emp}} - \tau^- P_{\text{emp}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}
\end{flalign*}
Note that $e^{-1} \leq P_{\text{est}}, P_{\text{est}}^- \leq e$. We extend the proofs of Lemma A.2 and Theorem 3 from \citep{chuang2021debiased} to get results for DebiasedPos Loss.

\begin{enumerate}[leftmargin=*]
    \item
\begin{flalign*}
    &\Delta_1 := |A_2 - A_1| = \bigg| -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} + \log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-} \bigg|.&\\
    &\forall \varepsilon > 0: \mathbb{P} (\Delta_1 \geq \varepsilon) = \mathbb{P} \bigg(\bigg|\log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-\}\bigg| \geq \varepsilon \bigg) = &\\
    & = \mathbb{P} \bigg(\log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-\} \geq \varepsilon \bigg) + &\\
    & = \mathbb{P} \bigg(\log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-\} \leq -\varepsilon \bigg)
\end{flalign*}

The first term can be bounded as:
\begin{flalign*}
    &\mathbb{P} \bigg(\log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-\} \geq \varepsilon \bigg) =&\\
    &= \mathbb{P} \bigg(\log \frac{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-} \geq \varepsilon \bigg)
        \leq
    \mathbb{P} \bigg(\frac{(N \tau^+ - \tau^-)(P_{\text{emp}}^- - P_{\text{est}}^-)}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-} \geq \varepsilon \bigg)=&\\
    &= \mathbb{P} \bigg(P_{\text{emp}}^- - P_{\text{est}}^- \geq \varepsilon \bigg\{\frac{1}{N \tau^+ - \tau^-} P_{\text{est}} + P_{\text{est}}^- \bigg\}\bigg)
        \leq
    \mathbb{P} (P_{\text{emp}}^- - P_{\text{est}}^- \geq \varepsilon e^{-1})&\\
\end{flalign*}

Here we used the fact that $\log x \leq x - 1$ for $x > 0$, and $1/(N \tau^+ - \tau^-) P_{\text{est}} + P_{\text{est}}^- \geq e^{-1}$. Likewise bounding the second term:
$$\mathbb{P} \bigg(\log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{est}}^-\} \leq -\varepsilon \bigg) \leq \mathbb{P} (P_{\text{est}}^- - P_{\text{emp}}^- \geq \varepsilon e^{-1})$$

and applying Hoeffding’s inequality, we have:
\begin{flalign*}
    &\mathbb{P} (\Delta_1 \geq \varepsilon) \leq \mathbb{P} (|P_{\text{emp}}^- - P_{\text{est}}^-| \geq \varepsilon e^{-1}) = \mathbb{P} \bigg(\bigg|\frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} - \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^Tf(\textbf{x}^-)}\bigg| \geq \varepsilon e^{-1}\bigg) \leq&\\
    &\leq 2 \exp\bigg(-\frac{2N \varepsilon^2 e^{-2}}{e - e^{-1}} \bigg) \leq 2 \exp\bigg(-\frac{2N \varepsilon^2}{e^3} \bigg) &\\
\end{flalign*}

Finally, we write the expectation of $\Delta_1$ as the integral of its tail probability
\begin{flalign*}
    &\big|\mathbb{E} A_2 - \mathbb{E} A_1 \big| \leq \mathbb{E} \Delta_1 \leq \int_0^\infty 2 \exp\bigg(-\frac{2N \varepsilon^2}{e^3} \bigg) d \varepsilon = \sqrt{\frac{\pi}{2N}} e^{3/2}.
\end{flalign*}

    \item
\begin{flalign*}
    &\Delta_2 := |A_3 - A_2| = \bigg| -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} + \log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} \bigg|.&\\
    &\forall \varepsilon > 0: \mathbb{P} (\Delta_2 \geq \varepsilon) = \mathbb{P} \bigg(\bigg|\log \{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\}\bigg| \geq \varepsilon \bigg)&\\
\end{flalign*}

Like before, we split the probability into sum of two terms, and bound the first term as:
\begin{flalign*}
    &\mathbb{P} \bigg(\log \{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} - \log \{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\} \geq \varepsilon \bigg) =&\\
    &= \mathbb{P} \bigg(\log \frac{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} \geq \varepsilon \bigg)
        \leq
    \mathbb{P} \bigg(\frac{P_{\text{emp}} - P_{\text{est}}}{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} \geq \varepsilon \bigg)=&\\
    &= \mathbb{P} \bigg(P_{\text{emp}} - P_{\text{est}} \geq \varepsilon \big\{P_{\text{est}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-\big\}\bigg)
        \leq
    \mathbb{P} (P_{\text{emp}} - P_{\text{est}} \geq \varepsilon e^{-1})&\\
\end{flalign*}

This holds because with $N$ big enough, $N \tau^+ \geq \tau^-$, and thus $1/(N \tau^+ - \tau^-) P_{\text{est}} + P_{\text{est}}^- \geq e^{-1}$.
\begin{flalign*}
    &\mathbb{P} (\Delta_2 \geq \varepsilon) \leq \mathbb{P} (|P_{\text{emp}} - P_{\text{est}}| \geq \varepsilon e^{-1}) = &\\
    &= \mathbb{P} \bigg(\bigg| \frac{1}{N+2} \bigg(\sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} + e^{f(\textbf{x})^T f(\textbf{v})} + e^{f(\textbf{x})^T f(\textbf{x})}\bigg) - \mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^Tf(\textbf{x}')} \bigg| \geq \varepsilon e^{-1}\bigg) \leq&\\
    &\leq 2 \exp\bigg(-\frac{2(N+2) \varepsilon^2}{e^3} \bigg) &\\
    &\big|\mathbb{E} A_3 - \mathbb{E} A_2 \big| \leq \mathbb{E} \Delta_2 \leq \int_0^\infty 2 \exp\bigg(-\frac{2(N+2) \varepsilon^2}{e^3} \bigg) d \varepsilon = \sqrt{\frac{\pi}{2(N+2)}} e^{3/2}.
\end{flalign*}

    \item
\begin{flalign*}
    &\Delta_3 := |A_3 - A_4| = \bigg| -\log \frac{P_{\text{est}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} + \log \frac{P_{\text{emp}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}\bigg|.&\\
    &\forall \varepsilon > 0: \mathbb{P} (\Delta_3 \geq \varepsilon) = \mathbb{P} \bigg(\bigg|\log \{P_{\text{emp}} - \tau^- P_{\text{est}}^-\} - \log \{P_{\text{est}} - \tau^- P_{\text{est}}^-\}\bigg| \geq \varepsilon \bigg)&\\
\end{flalign*}

Like before, we split the probability into sum of two terms, and bound the first term as:
\begin{flalign*}
    &\mathbb{P} \bigg(\log \{P_{\text{emp}} - \tau^- P_{\text{est}}^-\} - \log \{P_{\text{est}} - \tau^- P_{\text{est}}^-\} \geq \varepsilon \bigg) =&\\
    &= \mathbb{P} \bigg(\log \frac{P_{\text{emp}} - \tau^- P_{\text{est}}^-}{P_{\text{est}} - \tau^- P_{\text{est}}^-} \geq \varepsilon \bigg)
        \leq
    \mathbb{P} \bigg(\frac{P_{\text{emp}} - P_{\text{est}}}{P_{\text{est}} - \tau^- P_{\text{est}}^-} \geq \varepsilon \bigg)=&\\
    &= \mathbb{P} \bigg(P_{\text{emp}} - P_{\text{est}} \geq \varepsilon \big\{P_{\text{est}} - \tau^-P_{\text{est}}^-\big\}\bigg)
        \leq
    \mathbb{P} (P_{\text{emp}} - P_{\text{est}} \geq \varepsilon \tau^+ e^{-1})&\\
\end{flalign*}

This time the bound is different, because $P_{\text{est}} - \tau^-P_{\text{est}}^- = \tau^+ P_{\text{est}}^+ := \tau^+ \mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^Tf(\textbf{x}^+)} \geq \tau^+ e^{-1}$.
\begin{flalign*}
    &\mathbb{P} (\Delta_3 \geq \varepsilon) \leq \mathbb{P} (|P_{\text{emp}} - P_{\text{est}}| \geq \varepsilon \tau^+ e^{-1}) = &\\
    &= \mathbb{P} \bigg(\bigg| \frac{1}{N+2} \bigg(\sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} + e^{f(\textbf{x})^T f(\textbf{v})} + e^{f(\textbf{x})^T f(\textbf{x})}\bigg) - \mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^Tf(\textbf{x}')} \bigg| \geq \varepsilon \tau^+ e^{-1}\bigg) \leq&\\
    &\leq 2 \exp\bigg(-\frac{2(N+2) \varepsilon^2 (\tau^+)^2}{e^3} \bigg) &\\
    &\big|\mathbb{E} A_3 - \mathbb{E} A_4 \big| \leq \mathbb{E} \Delta_3 \leq \int_0^\infty 2 \exp\bigg(-\frac{2(N+2) \varepsilon^2 (\tau^+)^2}{e^3} \bigg) d \varepsilon = \frac{1}{\tau^+} \sqrt{\frac{\pi}{2(N+2)}} e^{3/2}.
\end{flalign*}

    \item
\begin{flalign*}
    &\Delta_4 := |A_5 - A_4| = \bigg| -\log \frac{P_{\text{emp}} - \tau^- P_{\text{emp}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} + \log \frac{P_{\text{emp}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-} \bigg|.&\\
    &\forall \varepsilon > 0: \mathbb{P} (\Delta_4 \geq \varepsilon) = \mathbb{P} \bigg(\bigg|\log \{P_{\text{emp}} - \tau^- P_{\text{est}}^-\} - \log \{P_{\text{emp}} - \tau^- P_{\text{emp}}^-\}\bigg| \geq \varepsilon \bigg)&\\
\end{flalign*}

Like before, we split the probability into sum of two terms, and bound the first term as:
\begin{flalign*}
    &\mathbb{P} \bigg(\log \{P_{\text{emp}} - \tau^- P_{\text{est}}^-\} - \log \{P_{\text{emp}} - \tau^- P_{\text{emp}}^-\} \geq \varepsilon \bigg) =&\\
    &= \mathbb{P} \bigg(\log \frac{P_{\text{emp}} - \tau^- P_{\text{est}}^-}{P_{\text{emp}} - \tau^- P_{\text{emp}}^-} \geq \varepsilon \bigg)
        \leq
    \mathbb{P} \bigg(\frac{\tau^-(P_{\text{emp}}^- - P_{\text{est}}^-)}{P_{\text{emp}} - \tau^- P_{\text{emp}}^-} \geq \varepsilon \bigg)=&\\
    &= \mathbb{P} \bigg(P_{\text{emp}}^- - P_{\text{est}}^- \geq \varepsilon \bigg\{\frac{1}{\tau^-} P_{\text{emp}} - P_{\text{emp}}^-\bigg\}\bigg)
        \leq
    \mathbb{P} \bigg(P_{\text{emp}}^- - P_{\text{est}}^- \geq \varepsilon \bigg( \frac{\tau^+}{\tau^-} e^{-1} - \delta\bigg)\bigg)&\\
\end{flalign*}

To get a lower bound of the expression in braces for big $N$, we will take a slightly smaller value than its limit.
\begin{flalign*}
    &\frac{1}{\tau^-} P_{\text{emp}} - P_{\text{emp}}^- = \frac{1}{\tau^-} \frac{1}{N+2} \bigg(\sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} + e^{f(\textbf{x})^T f(\textbf{v})} + e^{f(\textbf{x})^T f(\textbf{x})}\bigg) - \frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} =&\\
    &= \bigg(\frac{1}{\tau^-}\frac{1}{N+2} - \frac{1}{N}\bigg)\bigg(\sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)}\bigg) + \frac{e^{f(\textbf{x})^T f(\textbf{v})} + e^{f(\textbf{x})^T f(\textbf{x})}}{\tau^-(N+2)} \xrightarrow[N \to \infty]{} \frac{e^{-1}}{\tau^-} - e^{-1} = \frac{\tau^+}{\tau^-} e^{-1}&\\
\end{flalign*}

\begin{flalign*}
    &\mathbb{P} (\Delta_4 \geq \varepsilon) \leq \mathbb{P} \bigg(|P_{\text{emp}}^- - P_{\text{est}}^-| \geq \varepsilon \bigg( \frac{\tau^+}{\tau^-} e^{-1} - \delta\bigg)\bigg) = &\\
    &= \mathbb{P} \bigg(\bigg| \frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} - \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^Tf(\textbf{x}^-)} \bigg| \geq \varepsilon \bigg( \frac{\tau^+}{\tau^-} e^{-1} - \delta\bigg)\bigg) \leq&\\
    &\leq 2 \exp\bigg(-\frac{2N \varepsilon^2 ( (\tau^+/ \tau^-) e^{-1} - \delta)^2}{e} \bigg) =: 2 \exp\bigg(-\frac{2N \varepsilon^2 ( \tau^+/\tau^- - \delta)^2}{e^3} \bigg)&\\
    &\big|\mathbb{E} A_5 - \mathbb{E} A_4 \big| \leq \mathbb{E} \Delta_4 \leq \int_0^\infty 2 \exp\bigg(-\frac{2N \varepsilon^2 ( \tau^+/\tau^- - \delta)^2}{e^3} \bigg) d \varepsilon =&\\
    &= \frac{1}{\tau^+/ \tau^- - \delta} \sqrt{\frac{\pi}{2N}} e^{3/2} =: \bigg(\frac{\tau^-}{\tau^+} - \delta\bigg) \sqrt{\frac{\pi}{2N}} e^{3/2}.&\\
\end{flalign*}

\end{enumerate}

The result follows by summing all four absolute differences:
\begin{flalign*}
    &\big|\tilde{L}_{\text{DebiasedPos}}^N (f) - L_{\text{DebiasedPos}}^N (f)\big| = \big|\mathbb{E} A_1 - \mathbb{E} A_5\big| \leq \big|\mathbb{E} A_1 - \mathbb{E} A_2\big| + \big|\mathbb{E} A_2 - \mathbb{E} A_3\big| + \big|\mathbb{E} A_3 - \mathbb{E} A_4\big| + \big|\mathbb{E} A_4 - \mathbb{E} A_5\big| \leq&\\
    &\leq \bigg[\bigg(1 + \frac{\tau^-}{\tau^+} + \delta\bigg) \sqrt{\frac{\pi}{2N}} + \bigg(1 + \frac{1}{\tau^+}\bigg) \sqrt{\frac{\pi}{2(N + 2)}}\bigg] e^{3/2}&\\
\end{flalign*}

\end{proof}

\end{document}

### 2023-04-24 paper/main.tex
latest edit

@@
         Roman Isachenko\\
+        Moscow Institute of Physics and Technology\\
+        Yandex\\
+	\texttt{roman.isachenko@phystech.edu} \\
 	   \\
@@
 \begin{abstract}
-	Recently there has been a renewed interest in using contrastive learning for self-supervised representation learning.
+	% Recently there has been a renewed interest in using contrastive learning for self-supervised representation learning.
 A popular method for learning representations without labels is to compare similar (positive) and dissimilar (negative) pairs of samples.
@@
 This will further improve availability of accurate models for classification in tasks where extensive labeling is expensive or inaccessible.
+
+Recently contrastive learning has regained popularity as a self-supervised representation learning technique.
+It involves comparing positive (similar) and negative (dissimilar) pairs of samples to learn representations without labels.
+However, false negative and false positive errors in sampling lead to the loss function bias.
+This paper analyzes various ways to eliminate these biases.
+Based on the fully-supervised case, we develop debiased contrastive models that account for same-label datapoints without requiring knowledge of true labels, and explore their properties.
+Using the debiased representations, we measure accuracy of predictions in the classification task.
+The experiments are carried out on the CIFAR10 dataset, demonstrating the applicability and robustness of the proposed method in scenarios where extensive labeling is expensive or not feasible.
 \end{abstract}
@@
 Representation learning has become increasingly popular in recent years due to its ability to learn meaningful representations from large amounts of data, without the need for manual feature engineering.
-A widely used solution for this problem is contrastive learning, a technique that uses the principle of contrasting samples against each other to learn attributes that are common between data classes and attributes that set apart a data class from another.
+A popular solution for this problem is contrastive learning, a technique that uses the principle of contrasting samples against each other to learn shared and distinctive attributes between different data classes.
 It encourages the representations of similar pairs $(\textbf{x}, \textbf{x}^+)$ to be close, and those of dissimilar pairs $(\textbf{x}, \textbf{x}^-)$ to be more orthogonal.
@@
 Contrastive learning has proven to be effective for eliciting visual representations: SimCLR \citep{Chen2020SimCLR} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods.
-Its authors developed NT-Xent loss (normalized temperature-scaled cross entropy loss), that uses cosine similarity instead of scalar products.
+Its authors developed NT-Xent loss (normalized temperature-scaled cross-entropy loss), that uses cosine similarity instead of scalar products.
 
-Due to the unavailability of actual labels during training, negative instances $\textbf{x}^-$ are often randomly selected from the training data.
+During training, the lack of true labels often necessitates the random selection of negative instances $\textbf{x}^-$ from the training data.
 However, this approach results in sampling bias, where $\textbf{x}^-$ may actually be similar to $\textbf{x}$.
-This bias causes a significant drop in performance.
+This bias leads to a significant drop in performance \citep{hong2021unbiased}.
 
 To address this issue, \citep{khosla2021supervised} extend the self-supervised batch contrastive approach to the fully-supervised setting, allowing to effectively leverage label information.
-According to the experiments, supervised contrastive loss outperforms the base cross entropy, but only by a small amount.
-It also outperforms the cross entropy on robustness benchmark (ImageNet-C, which applies common naturally occuring perturbations such as noise, blur and contrast changes to the ImageNet dataset) and is less sensitive to hyperparameter changes.
+The experiments demonstrate that the supervised contrastive loss outperforms the base cross-entropy, albeit only by a marginal degree.
+It also outperforms the cross-entropy on robustness benchmark (ImageNet-C, which applies common naturally occurring perturbations such as noise, blur and contrast changes to the ImageNet dataset) and is less sensitive to hyperparameter changes.
 
 As for the case of self-supervised learning, a debiased contrastive loss was proposed to mitigate the bias of negative sampling by \citep{chuang2021debiased}.
-Given $N$ negative and $M$ positive samples, $\mathcal{L}^{N, M}_{\text{Debiased}} (f)$ is the loss function that corrects for False Negative samples.
-To remain unsupervised in practice, samples are acquired from the data distribution and a positive distribution mimicked by data augmentations, which can contatin False Positives.
+Given $N$ negative and $M$ positive samples, $L^{N, M}_{\text{DebiasedNeg}} (f)$ is the loss function that corrects for False Negative samples.
+To remain unsupervised in practice, samples are acquired from the data distribution and a positive distribution mimicked by data augmentations, which can contain False Positive samples.
 
-This work aims to develop a new algorithm for debiased sampling in contrastive learning, which accounts for False Positive as well as False Negative samples.
+This work aims to develop a novel algorithm for debiased sampling in contrastive learning, which accounts for False Positive as well as False Negative samples.
 
@@
 
-In \citep{chuang2021debiased} the distribution over classes is denoted by $\rho(c)$, the joint distribution $p_{x,c}(\textbf{x}, c) = p(\textbf{x}|c)\rho(c)$.
-Let $h : X \rightarrow C$ be the function assigning the latent class labels.
+Let $\mathcal{X}$ be the object space, and let $p(x)$ be the data distribution over $\mathcal{X}$, then the target embedding is $f: \mathcal{X} \rightarrow \mathbb{R}^d$.
+In \citep{chuang2021debiased} the distribution over a set of the discrete latent classes $\mathcal{C}$ is denoted by $\rho(c)$, hence the joint distribution $p_{x,c}(\textbf{x}, c) = p(\textbf{x}|c)\rho(c)$.
+Let $h : \mathcal{X} \rightarrow \mathcal{C}$ be the function assigning the class labels.
 Then $p^+_x(\textbf{x}') = p(\textbf{x}'|h(\textbf{x}') = h(\textbf{x}))$ is the probability of observing $\textbf{x}'$ as a positive example for $\textbf{x}$ and $p^-_x(\textbf{x}') = p(\textbf{x}'|h(\textbf{x}') \neq h(\textbf{x}))$ the probability of a negative example.
 It is assumed that the class probabilities $\rho(\textbf{x}) = \tau^+$ are uniform, and $\tau^- = 1 - \tau^+$ is the probability of observing any different class.
-Then the  ``ideal``, or unbiased loss to optimize would be:
+Then the  ``ideal'', or unbiased loss to optimize would be:
 
@@
 \end{equation}
-However, estimating $p(\textbf{x}')$ and $p^+_x(\textbf{x}')$ empirically is computationally expensive, and the authors propose a cheaper way to estimate $p^-_x(\textbf{x}')$, using $N$ samples $\{\textbf{u}_i\}_{i=1}^N$ from $p$ and $M$ samples $\{\textbf{v}_i\}_{i=1}^M$ from $p_x^+$, with temperature $t$:
+However, estimating $p(\textbf{x}')$ and $p^+_x(\textbf{x}')$ empirically is computationally expensive, and the authors suggest a cost-effective approach for estimating $p^-_x(\textbf{x}')$, using $N$ samples $\{\textbf{u}_i\}_{i=1}^N$ from $p$ and $M$ samples $\{\textbf{v}_i\}_{i=1}^M$ from $p_x^+$, with temperature $t$:
 
 \begin{equation} \label{eq:6}
-L_{\text{DebiasedNeg}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p; \textbf{x}_+ \sim p_x^+,\\ \{\textbf{u}_i\}_{i=1}^N \sim p^N \\ \{\textbf{v}_i\}_{i=1}^N \sim {p_x^+}^M}}  \bigg[ -\log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)} }{e^{f(\textbf{x})^T f(\textbf{x}^+)} + N g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_i\}_{i=1}^M\big)} \bigg],
+L_{\text{DebiasedNeg}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p; \textbf{x}_+ \sim p_x^+,\\ \{\textbf{u}_i\}_{i=1}^N \sim p^N \\ \{\textbf{v}_j\}_{j=1}^M \sim {p_x^+}^M}}  \bigg[ -\log \frac{e^{f(\textbf{x})^T f(\textbf{x}^+)} }{e^{f(\textbf{x})^T f(\textbf{x}^+)} + N g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_j\}_{j=1}^M\big)} \bigg],
 \end{equation}
 
-where the empirical estimates of $p$ and $p_x^-$ are defined as
+where the empirical estimate of the term corresponding to $p_x^-$ is defined as
 \begin{equation} \label{eq:7}
-g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_i\}_{i=1}^M\big) = \max \bigg\{ \frac{1}{\tau^-}\bigg(\frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} - \tau^+ \frac{1}{M} \sum \limits_{i=1}^M e^{f(\textbf{x})^T f(\textbf{v}_i)}\bigg), e^{-1/t}\bigg\}.
+g\big(\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_j\}_{j=1}^M\big) = \max \bigg\{ \frac{1}{\tau^-}\bigg(\frac{1}{N} \sum \limits_{i=1}^N e^{f(\textbf{x})^T f(\textbf{u}_i)} - \tau^+ \frac{1}{M} \sum \limits_{j=1}^M e^{f(\textbf{x})^T f(\textbf{v}_j)}\bigg), e^{-1/t}\bigg\}.
 \end{equation}
 
-Using this loss on SimCLR framework allowed to achieve higher accuracy compared to the regular biased loss.
-With a similar theoretical inference, this work is aimed at developing a new algorithm, that would additionally focus on eliminating bias caused by False Positives.
-
-% \paragraph{\underline{Tasks:}} 
-% \begin{itemize}
-%   \item reproduce Debiased Contrastive Learning results
-%   \item develop novel debiased contrastive loss that accounts for False Positives.
-%   \item test the algorithm on CIFAR10 and MNIST datasets
-% \end{itemize}
+The implementation of this loss for the SimCLR framework resulted in improved accuracy compared to the regular biased loss.
+Building on this theoretical foundation, our study aims to develop an algorithm that specifically addresses the issue of eliminating False Positive bias.
 
 \subsection{False Positive Samples} \label{FP}
-False Positive samples can be encountered in contrastive learning due to excessive or inappropriate augmentation.
-Similar to Debiased Contrastive Learning, we propose to estimate the positive sampling distribution in the loss function.
+Excessive or inappropriate augmentation leads to False Positive samples in contrastive learning.
+Similar to Debiased Contrastive Learning, we propose estimating the positive sampling distribution within the loss function.
 
@@
 
-Contrary to Debiased Contrastive Learning, drawing samples from $p(\textbf{x}')$ will also include positive labelled datapoints, since some of them could be actually negative.
+Unlike Debiased Contrastive Learning, drawing samples from $p(\textbf{x}')$ will include positive labeled data points that may be, in fact, negative.
 
 \newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}[theorem]
 \newtheorem{lemma}[theorem]{Lemma}
@@
 
-Using \ref{eq:8} and linearity of the expectation we complete the proof:
+Using equation \ref{eq:8} and the linearity of the expectation we complete the proof:
 $$\mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^T f(\textbf{x}^+)} = \frac{1}{\tau^+} \big(\mathbb{E}_{\textbf{x}' \sim p} e^{f(\textbf{x})^T f(\textbf{x}')} - \tau^- \mathbb{E}_{\textbf{x}^- \sim p_x^-} e^{f(\textbf{x})^T f(\textbf{x}^-)}\big) = R.$$
@@
 
-We will denote the limit by $\tilde{L}_{\text{DebiasedPos}}^N$ and rewrite it, multiplying the nominator and the denominator of the under-logarithm expression by $\tau^+$:
+We will denote the limit by $\tilde{L}_{\text{DebiasedPos}}^N$ and rewrite it, multiplying the numerator and the denominator of the under-logarithm expression by $\tau^+$:
 
@@
 
-With finite N:
+With finite $N$:
 \begin{equation} \label{eq:12}
@@
 \label{thm:asympt}
-If $M = 1$, for any embedding f and any $\delta > 0$ exists large enough N, that
+If $M = 1$, for any embedding f and any $\delta > 0$ exists large enough $N$, that
 \begin{equation} \label{eq:15}
@@
 
-When the number of positive samples $M > 1$, similar to \citep{khosla2021supervised} we will aggregate losses using the summation over positive samples located outside of the log.
+When the number of positive samples $M > 1$, similar to \citep{khosla2021supervised}, we will aggregate losses using the summation over positive samples located outside of the log.
 In our case this means taking all possible pairs of same-class (positive) elements, and averaging all corresponding $L_{\text{DebiasedPos}}^N (f)$ losses.
-We will refer to this method as \textit{pos-grouping}.
+We will refer to this method as \textit{loss-combination}.
 
@@
 \begin{equation} \label{eq:16}
-L_{\text{DebiasedPos}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \{\textbf{u}_i\}_{i=1}^N \sim {p^-}^N \\ \{\textbf{v}_i\}_{i=1}^M\sim {p^+}^M}} \frac{1}{M} \sum\limits^M \bigg[-\log \frac{P_{\text{emp}} - \tau^- P_{\text{emp}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}\bigg]
+L_{\text{DebiasedPos}}^{N, M} (f) = \mathbb{E}_{\substack{\textbf{x} \sim p \\ \{\textbf{u}_i\}_{i=1}^N \sim {p^-}^N \\ \{\textbf{v}_j\}_{j=1}^M\sim {p^+}^M}} \frac{1}{M} \sum\limits^M \bigg[-\log \frac{P_{\text{emp}} - \tau^- P_{\text{emp}}^-}{P_{\text{emp}} + (N \tau^+ - \tau^-) P_{\text{emp}}^-}\bigg]
 \end{equation}
 
-Another way to aggregate positive samples is to do so inside the estimates $P_{\text{emp}}$, following the example of \citep{chuang2021debiased} which we will refer to as \textit{loss-combination}.
-Emperically, the experiments in Section \ref{pos_aggr} show that while \textit{pos-grouping} is computationally less expensive, \textit{loss-combination} yields better accuracy with larger M.
+Another way to aggregate positive samples is to do so inside the estimates $P_{\text{emp}}$, following the example of \citep{chuang2021debiased}, which we will refer to as \textit{pos-grouping}.
+Empirically, the experiments in Section \ref{pos_aggr} show that while \textit{pos-grouping} is computationally less expensive, \textit{loss-combination} yields better accuracy with larger $M$.
 
@@
 discrepancy between the image (image representation $f$) and its description (description representation $t$) in online shopping.
-To solve this problem, when calculating the losses for each picture, we propose to weigh the losses by the cosine similarity of the anchor image vector and the description text vector.
+To solve this problem, when calculating the losses for each picture, we propose to weight the losses by the cosine similarity of the anchor image vector and the description text vector.
 
@@
 
-The goal of the experiment is to compare Multi-class N-pair Loss \ref{eq:1} with Negative-Debiased Loss \ref{eq:6}.
+The goal of the experiment is to compare Multi-class N-pair Loss [\ref{eq:1}] with Negative-Debiased Loss [\ref{eq:6}].
 We use CIFAR10 \citep{krizhevsky2009learning} dataset, an image recognition dataset for image classification.
 It consists of 60000 32x32 color images in 10 classes.
-The baseline negative-debiased model is SimCLR \citep{Chen2020SimCLR} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{Diederik2014Adam}, learning rate 0.001, and batch size 512.
+The baseline Negative-Debiased model is SimCLR \citep{Chen2020SimCLR} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{Diederik2014Adam}, learning rate 0.001, and batch size 512.
 All models are trained for 50 epochs and evaluated by training a linear classifier after fixing the learned embedding.
@@
 
-To test the performance of the loss under the assumptions made in \ref{FP}, we artificially exclude False Negative errors via accessing ground truth labels during training, and discover that Positive-Debiased loss still outperforms Negative-Debiased loss [Figure \ref{fig:fig3}].
+To test the performance of the loss under the assumptions made in Section \ref{FP}, we artificially exclude False Negative errors via accessing ground truth labels during training, and discover that Positive-Debiased loss still outperforms Negative-Debiased loss [Figure \ref{fig:fig3}].
 
@@
 \includegraphics[width=1\textwidth]{figures/base_vs_M=2.png}
-\caption{Classification accuracy on CIFAR10 with an increased positive sample size M}
+\caption{Classification accuracy on CIFAR10 with an increased positive sample size $M$}
 \label{fig:fig5}
@@
 
-Additionally we test all three losses with increased positive sample size M [Figure \ref{fig:fig5}].
-Larger M leads to better estimate of the loss both for Positive- and Negative-Debiased losses, yielding better accuracy as a result.
+Additionally, we test all three losses with increased positive sample size $M$ [Figure \ref{fig:fig5}].
+Larger $M$ leads to better estimate of the loss both for Positive- and Negative-Debiased losses, yielding better accuracy as a result.
 It is worth noting that in all of the experiments Positive-Debiased loss achieves better accuracy in the earlier stages of training both for Top-1 and Top-5 scores.
@@
 \subsection{Positive Sample Aggregation} \label{pos_aggr}
-In this experiment we compare methods of aggregation over positive samples introduced in Section \ref{pos_aggr_theory}.
+In this experiment, we compare methods of aggregation over positive samples introduced in Section \ref{pos_aggr_theory}.
 We train SimCLR with the same parameters, except for setting batch size to 64 when testing $M=4$ for faster training.
@@
 \begin{table}
-    \caption{Aggregation over M positive samples for DebiasedPos}
+    \caption{Aggregation over $M$ positive samples for DebiasedPos}
     \centering
@@
         \toprule
-        Method  & M & bs    & Training time (min)   & acc@1 & acc@5\\
+        Method  & $M$ & bs    & Training time (min)   & acc@1 & acc@5\\
         \midrule
@@
 \begin{enumerate}
-    \item Positive-Debiased loss is based on the assumption of true negative samples, i.e.
-with no False Negative errors, however, it still improves accuracy of classification when this assumption does not hold.
-    \item Another implicit assumption we make is that the classes have uniform distribution, which is not true for many datasets, and it is an interesting direction of further reseach to additionally account for the biases in the class distribution.
+    \item Positive-Debiased loss is based on the assumption of true negative samples, i.e., with no False Negative errors; however, it still improves accuracy of classification when this assumption does not hold.
+    \item Another implicit assumption we make is that the classes have uniform distribution, which is not true for many datasets, and it is an interesting direction of further research to additionally account for the biases in the class distribution.
 \end{enumerate}
@@
 In this paper, we propose a new method for debiasing contrastive loss based on mitigating biases caused by False Positive errors.
-Theoretical analysis shows that computationally Positive-Debiased and baseline losses are equivalent, however, in a scenario with an increased False Positive rate our loss yields better accuracy, which shows that it is more robust.
-Additionally, Positive-Debiased loss is more dominant in the early epochs.
-We further explore 2 mehtods of aggregating loss over positive samples, and infer that \textit{pos-grouping} trains faster, while \textit{loss-combination} leads to better performance.
+Theoretical analysis shows that computationally Positive-Debiased and baseline losses are equivalent.
+However, in a scenario with an increased False Positive rate our loss yields better accuracy, which shows that it is more robust.
+Moreover, Positive-Debiased loss is more dominant in the early epochs.
+We further explore 2 methods of aggregating loss over positive samples, and infer that \textit{pos-grouping} trains faster, while \textit{loss-combination} leads to better performance.
 
@@
 
-Like before, we split the probability into sum of two terms, and bound the first term as:
+We split the probability into a sum of two terms, and bound the first term as:
 \begin{flalign*}
@@
 
-This holds because with $N$ big enough, $N \tau^+ \geq \tau^-$, and thus $1/(N \tau^+ - \tau^-) P_{\text{est}} + P_{\text{est}}^- \geq e^{-1}$.
+This holds because with $N$ large enough, $N \tau^+ \geq \tau^-$, and thus $1/(N \tau^+ - \tau^-) P_{\text{est}} + P_{\text{est}}^- \geq e^{-1}$.
 \begin{flalign*}
@@
 
-This time the bound is different, because $P_{\text{est}} - \tau^-P_{\text{est}}^- = \tau^+ P_{\text{est}}^+ := \tau^+ \mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^Tf(\textbf{x}^+)} \geq \tau^+ e^{-1}$.
+This time the bound is different because $P_{\text{est}} - \tau^-P_{\text{est}}^- = \tau^+ P_{\text{est}}^+ := \tau^+ \mathbb{E}_{\textbf{x}^+ \sim p_x^+} e^{f(\textbf{x})^Tf(\textbf{x}^+)} \geq \tau^+ e^{-1}$.
 \begin{flalign*}
@@
 
-To get a lower bound of the expression in braces for big $N$, we will take a slightly smaller value than its limit.
+To get a lower bound of the expression in braces for a large $N$, we will take a slightly smaller value than its limit.
 \begin{flalign*}

### 2023-05-03 slides/main.tex
cleanup + final talk

--- 2023-04-06 slides/template.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beamer Presentation
-% LaTeX Template
-% Version 2.0 (March 8, 2022)
-%
-% This template originates from:
-% https://www.LaTeXTemplates.com
-%
-% Author:
-% Vel (vel@latextemplates.com)
-%
-% License:
-% CC BY-NC-SA 4.0 (https://creativecommons.org/licenses/by-nc-sa/4.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
 \documentclass[
 	11pt, % Set the default font size, options include: 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt
-	%t, % Uncomment to vertically align all slide content to the top of the slide, rather than the default centered
-	%aspectratio=169, % Uncomment to set the aspect ratio to a 16:9 ratio which matches the aspect ratio of 1080p and 4K screens and projectors
+	t, % Uncomment to vertically align all slide content to the top of the slide, rather than the default centered
+	% aspectratio=169, % Uncomment to set the aspect ratio to a 16:9 ratio which matches the aspect ratio of 1080p and 4K screens and projectors
+        russian
 ]{beamer}
 
-\graphicspath{{Images/}{./}} % Specifies where to look for included images (trailing slash required)
-
+\graphicspath{{Images/}{./}}
 \usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule for better rules in tables
 \usepackage[russian]{babel}
-% \usepackage[english,russian]{babel}
 \usepackage[dvipsnames]{xcolor}
@@
 
-%----------------------------------------------------------------------------------------
-%	SELECT LAYOUT THEME
-%----------------------------------------------------------------------------------------
-
 % Beamer comes with a number of default layout themes which change the colors and layouts of slides. Below is a list of all themes available, uncomment each in turn to see what they look like.
 
-%\usetheme{default}
+% \usetheme{default}
 %\usetheme{AnnArbor}
@@
 %\usetheme{Szeged}
-%\usetheme{Warsaw}
-
-%----------------------------------------------------------------------------------------
-%	SELECT COLOR THEME
-%----------------------------------------------------------------------------------------
+% \usetheme{Warsaw}
 
@@
 
-%\usecolortheme{albatross}
-%\usecolortheme{beaver}
-%\usecolortheme{beetle}
-%\usecolortheme{crane}
-%\usecolortheme{dolphin}
-%\usecolortheme{dove}
-%\usecolortheme{fly}
-%\usecolortheme{lily}
-%\usecolortheme{monarca}
-%\usecolortheme{seagull}
-%\usecolortheme{seahorse}
-%\usecolortheme{spruce}
-%\usecolortheme{whale}
-%\usecolortheme{wolverine}
-
-%----------------------------------------------------------------------------------------
-%	SELECT FONT THEME & FONTS
-%----------------------------------------------------------------------------------------
-
-% Beamer comes with several font themes to easily change the fonts used in various parts of the presentation.
-Review the comments beside each one to decide if you would like to use it.
-Note that additional options can be specified for several of these font themes, consult the beamer documentation for more information.
+% \usecolortheme{spruce}
+% \usecolortheme{whale}
 
@@
 %\usefonttheme{structuresmallcapsserif} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in small caps serif
-
-%------------------------------------------------
-
-%\usepackage{mathptmx} % Use the Times font for serif text
 \usepackage{palatino} % Use the Palatino font for serif text
-
-%\usepackage{helvet} % Use the Helvetica font for sans serif text
 \usepackage[default]{opensans} % Use the Open Sans font for sans serif text
-%\usepackage[default]{FiraSans} % Use the Fira Sans font for sans serif text
-%\usepackage[default]{lato} % Use the Lato font for sans serif text
 \usepackage{mathrsfs}
 \usepackage{amsmath}
-%----------------------------------------------------------------------------------------
-%	SELECT INNER THEME
-%----------------------------------------------------------------------------------------
-
-% Inner themes change the styling of internal slide elements, for example: bullet points, blocks, bibliography entries, title pages, theorems, etc.
-Uncomment each theme in turn to see what changes it makes to your presentation.
-
-%\useinnertheme{default}
 \useinnertheme{circles}
-%\useinnertheme{rectangles}
-%\useinnertheme{rounded}
-%\useinnertheme{inmargin}
-
-%----------------------------------------------------------------------------------------
-%	SELECT OUTER THEME
-%----------------------------------------------------------------------------------------
-
-% Outer themes change the overall layout of slides, such as: header and footer lines, sidebars and slide titles. Uncomment each theme in turn to see what changes it makes to your presentation.
-
-%\useoutertheme{default}
-%\useoutertheme{infolines}
-%\useoutertheme{miniframes}
-%\useoutertheme{smoothbars}
-%\useoutertheme{sidebar}
-%\useoutertheme{split}
-%\useoutertheme{shadow}
-%\useoutertheme{tree}
-%\useoutertheme{smoothtree}
+\useoutertheme{default}
 
@@
 
-%----------------------------------------------------------------------------------------
-%	PRESENTATION INFORMATION
-%----------------------------------------------------------------------------------------
-
-\title[]{Анализ смещения распределений в контрастивном обучении} % The short title in the optional parameter appears at the bottom of every slide, the full title in the main parameter is only on the title page
-
-\author[Лидия Троешестова \and Роман Исаченко]{Лидия Троешестова \and Роман Исаченко} % Presenter name(s), the optional parameter can contain a shortened version to appear on the bottom of every slide, while the main parameter will appear on the title slide
-
-\institute[]{МФТИ} % Your institution, the optional parameter can be used for the institution shorthand and will appear on the bottom of every slide after author names, while the required parameter is used on the title slide and can include your email address or additional information on separate lines
-
-\date[\today]{\today} % Presentation date or conference/meeting name, the optional parameter can contain a shortened version to appear on the bottom of every slide, while the required parameter value is output to the title slide
-
-%----------------------------------------------------------------------------------------
+\title[]{Анализ смещения распределений в контрастивном обучении}
+\author[Лидия Троешестова \and Роман Исаченко]{Лидия Троешестова \and Роман Исаченко}
+\institute[]{Московский физико-технический институт}
+\date[\today]{\today}
 
 \begin{document}
-
-%----------------------------------------------------------------------------------------
-%	TITLE SLIDE
-%----------------------------------------------------------------------------------------
-
 \begin{frame}
@@
 
-%----------------------------------------------------------------------------------------
-%	PRESENTATION BODY SLIDES
-%----------------------------------------------------------------------------------------
-
-\setbeamerfont{frametitle}{size=\normalsize}
+\setbeamerfont{frametitle}{size=\small}
 %------------------------------------------------
@@
     \frametitle{Цель исследования}
-
-\begin{block}{Цель}
-Исследовать методы устранения смещения распределений $p_x^-$ и $p_x^+$ без использования меток соответствующих классов.
+\footnotesize
+\begin{block}{ \small Цель}
+Исследовать влияние смещения распределения $p_x^+$ в задаче построения представлений без учителя.
 \end{block}
 
-\begin{block}{Идея}
-В новой функции потерь использовать оценку распределения $p_x^+$ в предопложении верности негативных объектов.
+\begin{block}{ \small Проблема}
+Наличие смещения в распределениях классов приводит к некорректным представлениям объектов.
+\end{block}
+
+\begin{block}{ \small Идея}
+Учесть смещения распределений классов путем построения несмещенной функции потерь.
+% Построить несмещенную функцию потерь, используя оценку распределения $p_x^+$ в предположении верности негативных объектов.
 \end{block}
@@
 \begin{frame}
-    \frametitle{Постановка задачи}
-\small
-$p^+_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как позитивный объект для $\mathbf{x}$.
-
-$p^-_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как негативный объект для $\mathbf{x}$.
-
-$\tau^+$ — вероятность 1 класса; $\tau^- = 1 - \tau^+$ — вероятность любого другого класса $\implies p(\mathbf{x'}) = \tau^+ p_x^+(\mathbf{x'}) + \tau^-p_x^-(\mathbf{x'})$
+    \frametitle{Смещение распределений при построении представлений}
+\footnotesize
+\begin{itemize}
+    \item $p^+_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как позитивный объект для $\mathbf{x}$.
+    \item $p^-_x(\mathbf{x'})$ — вероятность взять $\mathbf{x'}$ как негативный объект для $\mathbf{x}$.
+    \item $\tau^+$ — вероятность 1 класса;
+    \item $\tau^- = 1 - \tau^+$ — вероятность любого другого класса
+    \item $p(\mathbf{x'}) = \tau^+ p_x^+(\mathbf{x'}) + \tau^-p_x^-(\mathbf{x'})$
+\end{itemize}
 
@@
 \begin{frame}
-    \frametitle{Решение}
+    \frametitle{Учет ошибок первого рода}
 \scriptsize
@@
 \begin{theorem}
-Для произвольного эмбеддинга f и произвольного $\delta > 0$ существует достаточно большое $N$, что
+Для произвольного представления f и произвольного $\delta > 0$ существует достаточно большое $N$, что
 \begin{equation*}
@@
 \end{frame}
-%------------------------------------------------
-
-% \begin{frame}
-% 	\frametitle{Debiased Contrastive Learning}
-%  \framesubtitle{Ching-Yao Chuang et al., NeurIPS 2020} % Optional subtitle
-
-% \scriptsize Минимизация $L_{\text{Debiased}}$ соответствует минимизации верхней границы $L_{\text{Unbiased}}$ в задаче обучения с учителем.
-
-% \begin{equation*} \tiny
-% L_{\text{Debiased}}^{N, M} (f) = \mathbb{E}_{\substack{\mathbf{x} \sim p; \mathbf{x}_+ \sim p_x^+,\\ \{\mathbf{u_i}\}_{i=1}^N \sim p^N \\ \{\mathbf{v_i}\}_{i=1}^M \sim p_x^+^M}}  \bigg[ -\log \frac{e^{f(\mathbf{x})^T f(\mathbf{x}^+)}}{e^{f(\mathbf{x})^T f(\mathbf{x}^+)} + N g\big(\mathbf{x}, \{\mathbf{u_i}\}_{i=1}^N, \{\mathbf{v_i}\}_{i=1}^M\big)} \bigg],
-% \end{equation*}
-
-% \begin{equation*} \tiny 
-% g\bigg(\mathbf{x}, \{\mathbf{u_i}\}_{i=1}^N, \{\mathbf{v_i}\}_{i=1}^M\bigg) = \max \bigg\{ \frac{1}{\tau^-}\bigg(\frac{1}{N} \sum \limits_{i=1}^N e^{f(\mathbf{x})^T f(\mathbf{u_i})} - \tau^+ \frac{1}{M} \sum \limits_{i=1}^M e^{f(\mathbf{x})^T f(\mathbf{v_i})}\bigg), e^{-1/t}\bigg\}
-% \end{equation*}
-
-% \begin{figure}
-%     \includegraphics[width=0.7\linewidth]{Images/tsne.png}
-% \end{figure}
-
-% \end{frame}
-
 %------------------------------------------------
@@
 \begin{itemize}
-    \item $\mathcal{T} -$ семейство аугментаций (color distortion, gaussian blur) 
+    \item $\mathcal{T} -$ семейство аугментаций (color distortion, Gaussian blur) 
     \item Семплируются 2 аугментации $t, t' \sim \mathcal{T}$, применяются к каждому объекту.
-    \item Обучаем сеть-энкодер $f(\cdot)$ и MLP сеть-проекцию $g(\cdot)$, максимизируя соответсвтие представлений.
+    \item Обучаем сеть-энкодер $f(\cdot)$ и MLP сеть-проекцию $g(\cdot)$, максимизируя соответствие представлений.
 \end{itemize}
 \end{frame}
-
 %------------------------------------------------
@@
 \begin{frame}
-	\frametitle{Эксперимент с удалением FN и с добавлением FP}
+	\frametitle{Вычислительный эксперимент}
+\scriptsize
+
+\begin{block}{ \small Цель эксперимента}
+Сравнить качество представлений при использовании 3 функций потерь: Contrastive, DebiasedNeg и DebiasedPos. 
+\end{block}
+
+\begin{block}{ \small Обучение}
+\begin{itemize}
+    \item Обучаем SimCLR с энкодером ResNet-18 и оптимизатором Adam, 50 эпох с размером батча 512.
+    \item Фиксируем выученные представления изображений из CIFAR10 на обучающей выборке.
+\end{itemize}
+\end{block}
+
+\begin{block}{ \small Валидация}
+На тестовой выборке классифицируем объект, применяя top-K к банку представлений, считаем top-1 и top-5 accuracy.
+\end{block}
+
+\end{frame}
+%------------------------------------------------
+
+\begin{frame}
+    \frametitle{\small Эксперимент с удалением FN, эксперимент с добавлением FP}
+
+\scriptsize
+Функция потерь DebiasedPos устойчива к шуму и имеет превосходство на ранних эпохах.
+
 \begin{figure}
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{Images/base_vs_dropfn.png}
+  \includegraphics[width=0.87\linewidth]{Images/base_vs_dropfn.png}
   \label{fig:sub1}
 \end{subfigure}%
+
 \begin{subfigure}
   \centering
-  \includegraphics[width=0.95\linewidth]{Images/base_vs_noise.png}
+  \hspace*{+0.22cm}\includegraphics[width=0.87\linewidth]{Images/base_vs_noise.png}
   \label{fig:sub2}
@@
 \begin{frame}
-	\frametitle{Эксперимент с увеличением кол-ва положительных семплов}
+	\frametitle{Увеличение кол-ва позитивных объектов $M$}
+\scriptsize
+Способы агрегации по $M$:
+\begin{enumerate}
+    \item \textit{pos-grouping}: внутри оценки $\color{violet}P_{\text{emp}} (\textbf{x}, \{\textbf{u}_i\}_{i=1}^N, \{\textbf{v}_j\}_{j=1}^M)$ вместо $h(\textbf{x}, \textbf{v})$ используем $\frac{1}{M} \sum \limits_{j=1}^M h(\textbf{x}, \textbf{v}_j)$.
+    \item \textit{loss-combination}: для каждой пары позитивных объектов считаем $L_{\text{DebiasedPos}}^N (f)$ и берем среднее по всем значениям функции потерь.
+
+\end{enumerate}
+
 \begin{figure}
@@
   \centering
-  \includegraphics[width=1\linewidth]{Images/base_vs_M=2.png}
+  \includegraphics[width=1\linewidth]{Images/pg_vs_lc.png}
+  \label{fig:sub1}
+\end{subfigure}
+\label{fig:test}
+\end{figure}
+
+\scriptsize При увеличении $M$ точность значительно возрастает, причем \textit{loss-combination} лучше, чем \textit{pos-grouping}.
+
+\end{frame}
+
+%------------------------------------------------
+
+\begin{frame}
+	\frametitle{Увеличение кол-ва позитивных объектов $M$}
+\scriptsize
+\begin{table}
+    \tiny
+    \centering
+    \begin{tabular}{llllll}
+        \toprule
+        Method  & M & bs    & Training time (min)   & acc@1 & acc@5\\
+        \midrule
+        -   & 1 & 512   & 122.85    & 76.76 & 98.61\\
+        \midrule
+        pos-grouping    & 2 & 512   & \textbf{181.87}   & 80.1  & 98.86\\
+        loss-combination    & 2 & 512   & 192.82    & \textbf{80.5} & \textbf{98.93}\\
+        \midrule
+        pos-grouping    & 4 & 64    & \textbf{278.88}   & 81.75 & 99.01\\
+        loss-combination    & 4 & 64    & 308.25    & \textbf{82.44} & \textbf{99.25}\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:table}
+\end{table}
+
+\scriptsize Есть trade-off между производительностью и точностью.
+
+\begin{figure}
+\centering
+\begin{subfigure}
+  \centering
+  \includegraphics[width=1\linewidth]{Images/pg_vs_lc.png}
   \label{fig:sub1}
@@
 \end{frame}
+
 %------------------------------------------------
@@
 
-\scriptisize
-
 \begin{itemize}
-    \item Лосс с устраненным FP-смещением добивается точности, сопоставимой с DebiasedNeg, и имеет превосходство на ранних эпохах.
-    \item В случае удаления ошибок FN функции потерь DebiasedPos и DebiasedNeg имеют сопоставимые точности.
-    \item При сильном зашумлении датасета, когда FP-rate увеличен, DebiasedPos имеет большое преимущество в точности над DebiasedNeg.
-    \item При увеличенном кол-ве положитлеьных семплов точность для всех рассмотренных функций потерь возрастает.
+\small
+    \item При верности предположения о правильности негативных объектов функция потерь DebiasedPos работает корректно.
+    \vfill
+    \item DebiasedPos более устойчив к шумным датасетам: когда увеличена доля ошибок I рода, DebiasedPos имеет большое преимущество в точности.
+    \vfill
+    \item При увеличенном кол-ве позитивных объектов точность для всех рассмотренных функций потерь возрастает.
+    \vfill
+    \item Способ агрегации \textit{pos-grouping} менее затратный, а \textit{loss-combination} более точный.
 \end{itemize}
@@
 
-%------------------------------------------------------------------------------
-%	CLOSING SLIDE
-%------------------------------------------------------------------------------
-
-% \begin{frame}[plain] % The optional argument 'plain' hides the headline and footline
-% 	\begin{center}	
-% 		\bigskip\bigskip % Vertical whitespace
-% 		{\LARGE Спасибо за внимание!}
-% 	\end{center}
-% \end{frame}
-
-%----------------------------------------------------------------------------------------
-
 \end{document}

### 2023-05-03 code/datasets.py

@@ -1,5 +1,4 @@
 import typing
-import logging
 
 import torch
 import numpy as np

### 2023-05-03 code/main.py

@@ -1,6 +1,5 @@
 import argparse
 import os
-import itertools
 import logging
 import time
 
@@ -210,7 +209,7 @@ def main(dataset: str, loss: str, root: str, batch_size: int, model_arch, *, cud
 
     step = 0
     for epoch in tqdm(range(1, epochs + 1)):
-        train_loss, step = train(model, train_loader, loss_criterion, optimizer, batch_size,
+        _, step = train(model, train_loader, loss_criterion, optimizer, batch_size,
                                  cuda=cuda, writer=writer, m_agg_mode=m_agg_mode, step=step)
         if epoch % 1 == 0:
             test_acc_1, test_acc_5 = test(model, memory_loader, test_loader, cuda=cuda, class_cnt=c, top_k=top_k,

### 2023-05-03 paper/main.tex
latest edit
edit
citations check

@@
 
-\title{Mitigating distributional biases in contrastive learning}
+\title{Mitigating Distributional Biases in Contrastive Learning}
 
@@
 
-\renewcommand{\shorttitle}{\textit{arXiv} Template}
+% \renewcommand{\shorttitle}{\textit{arXiv} Template}
 
@@
 
-The basic metric for supervised similarity is \textit{triplet loss} \citep{schroff2015facenet}, where \textbf{x} is called \textit{anchor}, \textbf{x}$^+$ is a sample with the same label as anchor, and \textbf{x}$^-$ is a sample with the different label than anchor.
+The basic metric for supervised similarity is \textit{triplet loss} \citep{Schroff_2015}, where \textbf{x} is called \textit{anchor}, \textbf{x}$^+$ is a sample with the same label as anchor, and \textbf{x}$^-$ is a sample with the different label than anchor.
 An improvement of triplet loss is \textit{Multi-class N-pair Loss} invented by \citep{NIPS2016_6b180037}:
@@
 
-As for the case of self-supervised learning, a debiased contrastive loss was proposed to mitigate the bias of negative sampling by \citep{chuang2021debiased}.
+As for the case of self-supervised learning, a debiased contrastive loss was proposed to mitigate the bias of negative sampling by \citep{chuang2020debiased}.
 Given $N$ negative and $M$ positive samples, $L^{N, M}_{\text{DebiasedNeg}} (f)$ is the loss function that corrects for False Negative samples.
@@
 It consists of 60000 32x32 color images in 10 classes.
-The baseline Negative-Debiased model is SimCLR \citep{Chen2020SimCLR} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{Diederik2014Adam}, learning rate 0.001, and batch size 512.
+The baseline Negative-Debiased model is SimCLR \citep{chen2020simple} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{kingma2017adam}, learning rate 0.001, and batch size 512.
 All models are trained for 50 epochs and evaluated by training a linear classifier after fixing the learned embedding.

### 2023-05-03 paper/main.tex
citations check

@@
 
-Contrastive learning has proven to be effective for eliciting visual representations: SimCLR \citep{Chen2020SimCLR} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods.
+Contrastive learning has proven to be effective for eliciting visual representations: SimCLR \citep{chen2020simple} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods.
 Its authors developed NT-Xent loss (normalized temperature-scaled cross-entropy loss), that uses cosine similarity instead of scalar products.
@@
 However, this approach results in sampling bias, where $\textbf{x}^-$ may actually be similar to $\textbf{x}$.
-This bias leads to a significant drop in performance \citep{hong2021unbiased}.
+This bias leads to a significant drop in performance \citep{NEURIPS2021_de8aa43e}.
 

### 2023-05-03 paper/main.tex
citations check

@@
 \begin{abstract}
-	% Recently there has been a renewed interest in using contrastive learning for self-supervised representation learning.
-A popular method for learning representations without labels is to compare similar (positive) and dissimilar (negative) pairs of samples.
-However, negative samples are often chosen randomly, which means that they could be of the same class.
-This paper analyzes various ways to eliminate these biases.
-Based on the fully-supervised case, we develop debiased contrastive models that account for same-label datapoints without requiring knowledge of true labels, and explore their properties.
-The experiments are performed on CIFAR10 dataset.
-This will further improve availability of accurate models for classification in tasks where extensive labeling is expensive or inaccessible.
-
 Recently contrastive learning has regained popularity as a self-supervised representation learning technique.
@@
 
-The basic metric for supervised similarity is \textit{triplet loss} \citep{Schroff_2015}, where \textbf{x} is called \textit{anchor}, \textbf{x}$^+$ is a sample with the same label as anchor, and \textbf{x}$^-$ is a sample with the different label than anchor.
+The basic metric for supervised similarity is \textit{triplet loss} \citep{Schroff_2015}, where \textbf{x} is called \textit{anchor}, \textbf{x}$^+$ is a sample with the same label as anchor, and \textbf{x}$^-$ is a sample with a different label than anchor.
 An improvement of triplet loss is \textit{Multi-class N-pair Loss} invented by \citep{NIPS2016_6b180037}:
@@
 Let $\mathcal{X}$ be the object space, and let $p(x)$ be the data distribution over $\mathcal{X}$, then the target embedding is $f: \mathcal{X} \rightarrow \mathbb{R}^d$.
-In \citep{chuang2021debiased} the distribution over a set of the discrete latent classes $\mathcal{C}$ is denoted by $\rho(c)$, hence the joint distribution $p_{x,c}(\textbf{x}, c) = p(\textbf{x}|c)\rho(c)$.
+In \citep{chuang2020debiased} the distribution over a set of the discrete latent classes $\mathcal{C}$ is denoted by $\rho(c)$, hence the joint distribution $p_{x,c}(\textbf{x}, c) = p(\textbf{x}|c)\rho(c)$.
 Let $h : \mathcal{X} \rightarrow \mathcal{C}$ be the function assigning the class labels.
@@
 
-Another way to aggregate positive samples is to do so inside the estimates $P_{\text{emp}}$, following the example of \citep{chuang2021debiased}, which we will refer to as \textit{pos-grouping}.
+Another way to aggregate positive samples is to do so inside the estimates $P_{\text{emp}}$, following the example of \citep{chuang2020debiased}, which we will refer to as \textit{pos-grouping}.
 Empirically, the experiments in Section \ref{pos_aggr} show that while \textit{pos-grouping} is computationally less expensive, \textit{loss-combination} yields better accuracy with larger $M$.
@@
 It consists of 60000 32x32 color images in 10 classes.
-The baseline Negative-Debiased model is SimCLR \citep{chen2020simple} with ResNet-18 \citep{7780459} as the encoder architecture, with the Adam optimizer \citep{kingma2017adam}, learning rate 0.001, and batch size 512.
+The baseline Negative-Debiased model is SimCLR \citep{chen2020simple} with ResNet-18 \citep{he2015deep} as the encoder architecture, with the Adam optimizer \citep{kingma2017adam}, learning rate 0.001, and batch size 512.
 All models are trained for 50 epochs and evaluated by training a linear classifier after fixing the learned embedding.
@@
 \end{flalign*}
-Note that $e^{-1} \leq P_{\text{est}}, P_{\text{est}}^- \leq e$. We extend the proofs of Lemma A.2 and Theorem 3 from \citep{chuang2021debiased} to get results for DebiasedPos Loss.
+Note that $e^{-1} \leq P_{\text{est}}, P_{\text{est}}^- \leq e$. We extend the proofs of Lemma A.2 and Theorem 3 from \citep{chuang2020debiased} to get results for DebiasedPos Loss.
 
